---
title: "Initial analysis"
output: html_document
---


###### {-} 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(plotly)
library(htmltools)
library(sf)
library(here)
library(flextable)
library(ggrepel)
library(downloadthis)

interactive_colour <- function(n){
  color = c('#f46d43', "#FDAE6B", '#848484', '#e6e6e6')
  if(n==2){
    return(color[c(1,4)])
  }
  color[1:n]
}
color_binary <- c('#f46d43', '#e6e6e6')
palette = 'Oranges'
```

```{r data}
df <- readr::read_csv("/Users/jakepowell/Desktop/OME/Pilot/Data/Year 7_Survey Responses.csv") # pilot data location


group = c('Maths_in_Year_7' |> rep(6),
          'The_value_of_learning_maths'|> rep(7), 
          'Year_7_maths_lessons'|> rep(12), 
          'Learning_maths_outside_of_lessons'|> rep(4),
          'Your_confidence_in_maths'|> rep(12), 
          'Your_enjoyment_of_maths'|> rep(7),
          'Your_feelings_when_doing_maths'|> rep(6), 
          'What_is_mathsXX'|> rep(7))              

names(df)[6:ncol(df)] = paste0(group, '__', names(df)[6:ncol(df)]) 
names(df) = names(df) |> stringr::str_replace_all(pattern = ' ', replacement = '_')

df$`Group_/_class` = df$`Group_/_class` |> stringr::str_replace_all(pattern = '/','-')
school = df$Barcode_ID |> as.character() |>  substr(3, 5)
school_and_class = paste0(school,'__', df$`Group_/_class`)

pilot = data.frame(df[,1:4], school, school_and_class, df[5:ncol(df)])
for(i in 1:ncol(pilot)){
  pilot[,i][pilot[,i] == 'N-A'] = NA
  pilot[,i] = pilot[,i] |> stringr::str_to_sentence()
}


survey_data = pilot
survey_name = 'Year 7 pilot study'
report_kind = 'interactive'
```

```{css, echo = FALSE}
/* from https://ianlunn.github.io/Hover/ */
.hvr-sweep-to-left {
  display: inline-block;
  vertical-align: middle;
  -webkit-transform: perspective(1px) translateZ(0);
  transform: perspective(1px) translateZ(0);
  box-shadow: 0 0 1px rgba(0, 0, 0, 0);
  position: relative;
  -webkit-transition-property: color;
  transition-property: color;
  -webkit-transition-duration: 0.3s;
  transition-duration: 0.3s;
}
.hvr-sweep-to-left:before {
  content: "";
  position: absolute;
  z-index: -1;
  top: 0;
  left: 0;
  right: 0;
  bottom: 0;
  background: #f46d43;
  -webkit-transform: scaleX(0);
  transform: scaleX(0);
  -webkit-transform-origin: 100% 50%;
  transform-origin: 100% 50%;
  -webkit-transition-property: transform;
  transition-property: transform;
  -webkit-transition-duration: 0.3s;
  transition-duration: 0.3s;
  -webkit-transition-timing-function: ease-out;
  transition-timing-function: ease-out;
}
.hvr-sweep-to-left:hover, .hvr-sweep-to-left:focus, .hvr-sweep-to-left:active {
  color: white;
}
.hvr-sweep-to-left:hover:before, .hvr-sweep-to-left:focus:before, .hvr-sweep-to-left:active:before {
  -webkit-transform: scaleX(1);
  transform: scaleX(1);
}

.blackbox {
  padding: 1em;
  background: #FDAE6B;
  border: 2px solid #e6e6e6;
  border-radius: 10px;
}
.center {
  text-align: center;
}
```


*In this document we produce an initial automated analysis of the `r survey_name` survey.*

### Evaluate Participation

```{r, headline numbers}
no_rows = nrow(survey_data)
empty_index = which(is.na(survey_data$First_Name) & is.na(survey_data$Surname))
no_empty_surv = length(empty_index)
valid_percent = ((no_rows-no_empty_surv)/no_rows*100) |> round(1)

empty_surv_ID = survey_data$Barcode_ID[empty_index]

participation = survey_data$Tick_Appropriate_Box_Only_if_the_Child_did_not_take_the_Test
participation[is.na(participation)] = 'Participating'
if(no_empty_surv > 0){
  participation[empty_index] = 'Blank'
}

participation_table = participation |> table() |> data.frame()
participation_table$percent = (participation_table$Freq / sum(participation_table$Freq) *100) |> round(1)

if(no_empty_surv > 0){
participation_table$percent_no_blank = (participation_table$Freq / sum(participation_table$Freq[participation_table$participation != 'Blank']) *100) |> round(1)
participation_table$percent_no_blank[participation_table$participation == 'Blank'] = '-'
}

participation_table = participation_table[order(participation_table$Freq, decreasing = T),]
names(participation_table) = c('Participation', 'No responses', '%', '% (exc. blanks)')

table_to_show =  participation_table |> DT::datatable(extensions = 'Buttons',
                  rownames = FALSE,
                  options = list(scrollX = TRUE,
                                 pageLength = min(nrow(participation_table), 10),
                                 dom = 'Btp',
                                 buttons = c('copy', 'csv', 'excel', 'pdf'),
                                 paging=F,
                                 ordering=F))


fig <- participation_table |> plot_ly(labels = ~Participation,
                                      values = ~`No responses`,
                                      textposition = 'outside',
                                      textinfo = 'label+percent', title = 'Participation') |>
  add_pie(hole = 0.6) |>
  layout(title = "",  showlegend = F,
                      xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                      yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE)
         )


fig_no_blanks <- participation_table[participation_table$Participation !='Blank',] |> plot_ly(labels = ~Participation,
                                      values = ~`No responses`,
                                      textposition = 'outside',
                                      textinfo = 'label+percent') |>
  add_pie(hole = 0.6) |>
  layout(title = "",  showlegend = F,
                      xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                      yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE)
         )


```

`r if(no_empty_surv == 0){paste0("The survey data contains ", no_rows, " returned and scanned questionnaires, there are no blank surveys.")}`

`r if(no_empty_surv > 0){paste0("Out of the ", no_rows, " returned and scanned questionnaires, ", no_rows - no_empty_surv, " (", valid_percent, "%) include a pupil's name. The remaining ", no_empty_surv, " copies are  blank (those without first name and surname)")}`


`r if(no_empty_surv > 0){"<details>"}`
`r if(no_empty_surv > 0){"<summary>Barcodes of blank scanned questionnaires</summary>"}`
`r if(no_empty_surv > 0){paste0(empty_surv_ID, collapse = ', ')}`
`r if(no_empty_surv > 0){"</details>"}`

`r if(no_empty_surv > 0){""}`

Below we produce a chart and table summarising the participation of individuals across the survey.

##### {.tabset}

###### Chart

```{r, fig.fullwidth=TRUE, fig.dim = c(10, 4)}
fig
```

###### Table

```{r}
table_to_show
```
##### {-}

***

### Clean Data
### Evaluate Missingness and Distributions of Survey Item Responses
### Evaluate Dimensionality of Survey Items - Part 1 (Confirmatory Factor Analysis, CFA)
### Evaluate Dimensionality of Survey Items - Part 2 (Exploratory Graph Analysis, EGA)
### Explore Latent Groups (Latent Class Analysis, LCA)
### Item-wise School Comparison
### Scale-wise School Comparison
### Scale-wise Class Comparison
