---
title: "`r params$output_title`"
author: "`r params$output_author`"
date: "`r params$output_date`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
    toc_float: true
    code_folding: hide
    theme: cosmo
params:
  data_path: NULL
  dict_path: NULL
  dict_sheet: NULL
  est_chars_path: NULL
  est_chars_sheet: NULL
  est_char_vars: NULL
  est_char_types: NULL
  est_char_values: NULL
  est_char_statements: NULL
  output_title: NULL
  output_author: NULL
  output_date: NULL
  verbose: NULL
---


# Setup


## Preliminaries
```{r setup, include=FALSE, echo=FALSE}
#knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.height=4)
knitr::opts_chunk$set(fig.height=2.5)



```

```{r results='asis'}
if (!requireNamespace("cowplot", quietly=TRUE)){
  cat(report_notice(
    "Package 'cowplot' is not detected. It is needed to centre legends underneath plots with respect to the whole figure rather than just the plot area. Any section summary plots with long labels on the vertical axis might benefit a bit from  install.packages('cowplot') and then re-running.",
    level = "NOTE"))
}
```


```{r}
read_output <- survey_read_inputs(
  data_path = params$data_path,
  dict_path = params$dict_path,
  dict_sheet = params$dict_sheet,
  est_chars_path = params$est_chars_path,
  est_chars_sheet = params$est_chars_sheet,
  est_char_vars = params$est_char_vars,
  est_char_types = params$est_char_types,
  est_char_values = params$est_char_values,
  est_char_statements = params$est_char_statements
)


data <- read_output$data
dict <- read_output$dict
read_messages <- read_output$messages



#If verbose spit out the dictionary
if (params$verbose){
  print("VERBOSE MODE: here is the data dictionary")
  
  dict |>
    dplyr::mutate(dplyr::across(dplyr::where(is.character), # in character columns change newline (many possibilities) to double-space
                                ~ stringr::str_replace_all(., "\\r?\\n", "  "))) |>
    reactable::reactable(wrap=TRUE,
                         #columns = col_width,
                         showPageSizeOptions = TRUE,
                         pageSizeOptions = c(10,25,50,100),
                         defaultPageSize = 10,
                         fullWidth = FALSE,
                         striped=TRUE,
                         compact=TRUE,
                         filterable=TRUE,
                         resizable = TRUE)
  }

```



```{r results='asis'}
for (m in read_messages) {
  cat(report_notice(m$text, level = m$level))
}
```







```{r}
prep_out <- survey_data_prepare(data, dict)

data <- prep_out$data
validation_log <- prep_out$validation_log
validation_df <- prep_out$validation_df
prep_messages <- prep_out$messages
```


```{r}
if (params$verbose){
  print("VERBOSE MODE: here is the validation log (in tibble, not list, form)")

  validation_df |>
    dplyr::mutate(dplyr::across(dplyr::where(is.character), # in character columns change newline (many possibilities) to double-space
                                ~ stringr::str_replace_all(., "\\\\r\\\\n|\\r\\n|\\n|\\r", "  "))) |>
    reactable::reactable(wrap=TRUE,
                         #columns = col_width,
                         showPageSizeOptions = TRUE,
                         pageSizeOptions = c(10,25,50,100),
                         defaultPageSize = 10,
                         fullWidth = FALSE,
                         striped=TRUE,
                         compact=TRUE,
                         filterable=TRUE,
                         resizable = TRUE)
}
```

```{r}  
if (params$verbose){
  print("VERBOSE MODE: here is the str() of the data")

  data |> str(strict.width="cut")
}
```

```{r results='asis'}
for (m in prep_messages) {
  cat(report_notice(m$text, level = m$level))
}
```


## Substantive reporting starts here



Subsequent sectioning & content is as coded in the data dictionary document.

```{r}
percCutHoriz <- 2
percCutVert <- 2
```

Note that, for categorical variables, percentages on plots are suppressed if

- they are less than `r percCutHoriz`% on the horizontal-barred beginning-of-section summary plots and
- if less than `r percCutVert`% on the vertical-barred question and question-by-variable plots that are within subsections.

Also

- “Missing” counts include only truly missing (NA) values.
- “Non‑allowed” counts include values that are present but violate data dictionary or type constraints.


# Missingness

A short summary of section-by-section complete missingness.

- In this section, missing includes explicit non‑response codes (−999/−888/−777). In the validation summaries, these are treated separately (as "non-allowed").
- A section missing means that all questions in that section have no substantive response.


```{r}


# ------------------------------------------------------------------
# Section-level non-response summary
# A response is missing if it is NA OR one of the sentinel codes
# ------------------------------------------------------------------

missing_codes <- c(-999, -888, -777)

# Map sections to variables (excluding Grouping variables)
sec_vars <-
  validation_df |>
  dplyr::filter(
    !is.na(report_sec),
    report_sec != "Grouping variables"
  ) |>
  dplyr::arrange(report_sec, report_sec_ord) |>
  dplyr::summarise(
    vars = list(variable),
    .by = report_sec
  )

data_miss_flags <- data
n_resp <- nrow(data)

# ---- Base section-level flags: all_missing_sec_i ----
for (i in seq_len(nrow(sec_vars))) {
  
  v_all <- sec_vars$vars[[i]]
  v_present <- v_all[v_all %in% names(data)]
  nm <- paste0("all_missing_sec_", i)
  
  if (length(v_present) == 0) {
    # Section exists conceptually but has no variables in data
    data_miss_flags[[nm]] <- rep(FALSE, n_resp)
    next
  }
  
  m <- as.matrix(
    as.data.frame(
      lapply(
        data[v_present],
        function(x) is.na(x) | (x %in% missing_codes)
      )
    )
  )
  
  data_miss_flags[[nm]] <- rowSums(!m) == 0
}

# ---- Cumulative flags: all_missing_sec_i_on (ALWAYS created) ----
sec_cols <- paste0("all_missing_sec_", seq_len(nrow(sec_vars)))

for (i in seq_along(sec_cols)) {
  data_miss_flags[[paste0(sec_cols[i], "_on")]] <-
    Reduce(`&`, data_miss_flags[sec_cols[i:length(sec_cols)]])
}



# ---- Aggregate summary ----

miss_summary <-
  data_miss_flags |>
  dplyr::summarise(
    dplyr::across(
      dplyr::matches("^all_missing_sec_\\d+$"),
      ~ sum(.x),
      .names = "{.col}"
    ),
    dplyr::across(
      dplyr::matches("^all_missing_sec_\\d+_on$"),
      ~ sum(.x),
      .names = "{.col}"
    )
  ) |>
  tidyr::pivot_longer(
    tidyselect::everything(),
    names_to = "key",
    values_to = "num"
  ) |>
  dplyr::mutate(
    type = dplyr::if_else(
      grepl("_on$", key),
      "This section and after",
      "This section only"
    ),
    sec_index = readr::parse_number(key),
    sec = sec_vars$report_sec[sec_index],
    prop = num / n_resp
  )

```

```{r}

# ---- Table ----


miss_table <-
  miss_summary |>
  dplyr::select(sec, type, num, prop) |>
  tidyr::pivot_wider(
    names_from = type,
    values_from = c(num, prop)
  )



miss_table |>
  dplyr::mutate(
    `This section only – Proportion`
      = formatC(`prop_This section only`, digits = 3, format = "f"),
    `This section and after – Proportion`
      = formatC(`prop_This section and after`, digits = 3, format = "f")
  ) |>
  dplyr::select(
    Section = sec,
    `This section – N` = `num_This section only`,
    `This section – Proportion` = `This section only – Proportion`,
    `This section and after – N` = `num_This section and after`,
    `This section and after – Proportion` = `This section and after – Proportion`
  ) |>
  dplyr::bind_rows(
    tibble::tibble(
      Section = "Total",
      `This section – N` = n_resp,
      `This section – Proportion` = "--",
      `This section and after – N` = n_resp,
      `This section and after – Proportion` = "--"
    )
  ) |>
  kable_narrow()
```

```{r fig.height=6, fig.width=8}

# ---- Plot ----


miss_summary |>
  ggplot2::ggplot(
    ggplot2::aes(
      x = sec,
      y = prop,
      colour = type,
      group = type
    )
  ) +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::labs(
    x = NULL,
    y = "Proportion of respondents",
    colour = NULL,
    title = "Section-level complete non-response"
  ) +
  OMESurvey::theme_OME() +
  ggplot2::theme(
    axis.text.x = ggplot2::element_text(angle = 45, hjust = 1),
    legend.position = "right",
    legend.direction = "vertical"
  ) +
  ggplot2::scale_colour_discrete(
    labels = c(
      "This section only" = "This section\nonly",
      "This section and after" = "This section\nand after"
    )
  )

```




```{r}
emit_reactable <- function(df, ...) {
  child <- paste0(
    "```{r echo=FALSE}\n",
    "reactable::reactable(df, ...)\n",
    "```\n"
  )
  out <- knitr::knit_child(text = child, envir = environment(), quiet = TRUE)
  cat(out, sep = "\n")
}



freq_to_df <- function(freq, value_col = "Value", add_total = TRUE) {
  if (is.null(freq) || length(freq) == 0) return(NULL)

  df <- data.frame(
    setNames(list(names(freq)), value_col),
    Frequency = as.integer(freq),
    check.names = FALSE
  )

  if (isTRUE(add_total)) {
    df <- rbind(
      df,
      data.frame(
        setNames(list("Total"), value_col),
        Frequency = sum(df$Frequency, na.rm = TRUE),
        check.names = FALSE
      )
    )
  }

  df
}

remove_common_prefix <- function(x) {
  # find longest common prefix length
  pref_len <- 0L
  for (len in seq_len(min(nchar(x)))) {
    if (all(substr(x, 1, len) == substr(x[1], 1, len))) {
      pref_len <- len
    } else break
  }
  # build results
  if (pref_len == 0L) {
    list(prefix = "", remainder = x)
  } else {
    prefix <- substr(x[1], 1, pref_len)
    remainder <- substring(x, pref_len + 1)
    list(prefix = prefix, remainder = remainder)
  }
}



parse_report_sec <- function(sec) {
  s <- stringr::str_squish(as.character(sec))

  # Special case: Grouping variables (no prefix, always first)
  if (tolower(s) == "grouping variables") {
    return(list(
      display  = "Grouping variables",
      letter   = NA_character_,
      digit    = NA_integer_,
      code     = "GV",       # filename-safe short code
      sort_key = "0",        # always first
      file_key = "GV"
    ))
  }

  # General case:
  # Allow "A: Title" OR "A1: Title" where digit is 1-9
  m <- stringr::str_match(s, "^([A-Z])([1-9][0-9]?)?:\\s*(.+)$")
  if (!is.na(m[1, 1])) {
    letter <- m[1, 2]
    digit  <- if (!is.na(m[1, 3])) as.integer(m[1, 3]) else NA_integer_
    title  <- stringr::str_squish(m[1, 4])

    code <- if (is.na(digit)) letter else paste0(letter, digit)

    # sort: letter first, then digit (blank treated as 0, so "A:" comes before "A1:")
    digit_sort <- if (is.na(digit)) 0L else digit

    return(list(
      display  = paste0(code, ": ", title),
      letter   = letter,
      digit    = digit,
      code     = code,
      sort_key = sprintf("1_%s_%02d", letter, digit_sort),
      file_key = code
    ))
  }

  stop(
    "Invalid report_sec: '", s, "'. Allowed forms are:\n",
    "  - 'Grouping variables'\n",
    "  - 'X: Section title' where X is A–Z\n",
    "  - 'Xy: Section title' where X is A–Z and y is an integer 1–99 (no leading zero)"
  )
}

```

```{r results='asis', echo=FALSE, message=TRUE, fig.width=10, fig.height=6}

# order sections and ensure variables are ordered within each section

sections_raw <-
  validation_df |>
  dplyr::pull(report_sec) |>
  as.character() |>
  stats::na.omit() |>
  unique()

sec_meta <-
  tibble::tibble(sec_raw = sections_raw) |>
  dplyr::rowwise() |>
  dplyr::mutate(info = list(parse_report_sec(sec_raw))) |>
  dplyr::mutate(
    sec_display  = info$display,
    sec_letter   = info$letter,
    sec_sort_key = info$sort_key,
    sec_file_key = info$file_key
  ) |>
  dplyr::ungroup()




# Validate: section codes must be unique
# (e.g. "A", "A1", "B2", "GV" must not repeat)

dups <- sec_meta |>
  dplyr::group_by(sec_file_key) |>
  dplyr::filter(dplyr::n() > 1) |>
  dplyr::summarise(
    sections = paste(sprintf('"%s"', sec_display), collapse = ", "),
    .groups = "drop"
  )

if (nrow(dups) > 0) {
  msg <- paste(
    sprintf(
      'Key "%s" is used by sections: %s',
      dups$sec_file_key,
      dups$sections
    ),
    collapse = "\n"
  )
  
  stop(
    "Duplicate section codes (file keys) found:\n\n",
    msg,
    "\n\nEach code (before the colon in the section title) must be unique."
  )
}


# Order: Grouping variables first (sort_key "0"), then A, B, C...
sec_meta <- sec_meta |>
  dplyr::arrange(sec_sort_key)

sections <- sec_meta$sec_display





for(sec in sections){
  cat("\n\n# ", sec, "\n\n", sep = "")
  
  # catch potential file-naming issues
  sec_file_key <- sec_meta$sec_file_key[match(sec, sec_meta$sec_display)]
  if (is.na(sec_file_key)) stop("Could not determine file key for section: ", sec)
  
  
  vars_in_sec <-
    validation_df |>
    dplyr::filter(report_sec == sec) |>
    dplyr::arrange(report_sec_ord)
  
  vars <- vars_in_sec$variable
  #vars |> print()
  
  if(length(vars) == 0) next
  
  # vars_in_sec |>
  #   dplyr::select(type, allowed_specified, allowed) |> print()
  # 
  # vars_in_sec |>
  #   dplyr::select(type, allowed_specified, allowed) |>
  #   dplyr::distinct() |> print()
  
  
  #if not relevant, make a note that there's no summary plot, otherwise do it
  if (length(vars) == 1 || (vars_in_sec |>
                            dplyr::select(type, allowed_specified, allowed) |>
                            dplyr::distinct() |>
                            nrow()) != 1
  ) {
    cat(report_notice(
      "No section summary plot for this section.",
      level = "NOTE"
    ))
  } else {
    
    # prepare df and labels for plotting
    df_plot <- data |> dplyr::select(any_of(vars))
    labels_vec <-
      as.character(vars_in_sec$item_statement) |>
      setNames(vars_in_sec$variable)
    
    
    # get common prefix out of labels_vec to use as title
    labels_prefix <- remove_common_prefix(labels_vec)
    title_text <- labels_prefix$prefix
    labels_vec <- labels_prefix$remainder
    
    
    # determine section type
    sec_type <- vars_in_sec$type[[1]]
    #sec_type |> print()
    
    # compute dynamic height: 2 inches + 0.5 inch per variable
    fig_h <- (2 + 0.5 * length(vars))
    
    
    
    
    
    tryCatch({
      
      if (sec_type == "factor") {
        
        # use colo and ordering from the first variable
        colo <- (vars_in_sec |> dplyr::pull(colo) |> utils::head(1))[[1]]
        order_values <- (vars_in_sec |> dplyr::pull(order_values) |> utils::head(1))[[1]]
        
        # factor section summary plot
        n_cat <- nlevels(df_plot[[1]])
        fill_label_width <-
          if (n_cat <= 5) 20 else max(8, floor(100 / n_cat))
        
        res <- OMESurvey::summary_plot_stacked_bar(
          df_plot,
          labels_vec,
          percCut = percCutHoriz,
          colo = colo,
          order_values = order_values,
          titleText = title_text,
          NA_label = "Missing / Non-allowed",
          fill_label_width = fill_label_width
        ) |>
          OMESurvey::centre_legend_below(rel_heights = c(1, 0.8 / fig_h))
        
      } else if (sec_type == "numeric") {
        
        # numeric section summary plot
        res <- OMESurvey::summary_plot_boxplot(
          df_plot,
          labels_vec = labels_vec,
          titleText = title_text
        )
        
      } else {
        res <- NULL
      }
      
      
      if (inherits(res, "ggplot")) {
        # ensure fig path exists
        fig_dir <- knitr::opts_chunk$get("fig.path")
        if (is.null(fig_dir) || fig_dir == "") fig_dir <- "figure/"
        if (!dir.exists(fig_dir)) dir.create(fig_dir, recursive = TRUE)
        
        # create a unique, stable filename (using section name)
        fname <- file.path(
          fig_dir,
          sprintf("sec_%s_plot_overall.png", sec_file_key)
        )
        fname <- normalizePath(fname, winslash = "/", mustWork = FALSE)
        
        # save the plot and emit markdown image tag (for inside an as-is chunk)
        ggplot2::ggsave(fname, plot = res,
                        width = 8, height = fig_h,
                        units = "in", dpi = 300)
        
        cat(sprintf("![](%s){width=100%%}\n\n", fname))
      }
      
      invisible(res)
      
    }, error = function(e) {
      cat(report_notice(
        "Error drawing section plot: ", e$message
      ))
      invisible(NULL)
    })
  }
  
  
  
  
  # now emit per-variable summaries for this section
  for(i in seq_len(nrow(vars_in_sec))){
    v <- vars_in_sec[i, ] # the current row of vars_in_sec, in turn from validation_df. So has
    var <- v$variable
    label <- ifelse(is.na(v$item_statement) || v$item_statement == "", var, v$item_statement)
    
    
    # select grouping variables for this report; excluding the variable itself if appropriate
    grouping_vars <-
      dict |>
      dplyr::filter(grouping_var == TRUE) |>
      dplyr::pull(variable_name) |>
      setdiff(var)   # ensure the target var is not duplicated
    
    # ensure grouping_vars exist in data (safe select)
    grouping_vars <- grouping_vars[grouping_vars %in% names(data)]
    
    
    # extract condition for this variable
    cond <-
      dict |>
      dplyr::filter(variable_name == vars[i]) |>
      dplyr::pull(condition)
    
    # parse condition safely with tryCatch and friendly error message
    cond_expr <- NULL
    if (!is.na(cond) && nzchar(cond)) {
      cond_expr <- tryCatch(
        rlang::parse_expr(cond),
        error = function(e) {
          cat(report_notice(
            "Error parsing filter condition: ", e$message
          ))
          NULL
        }
      )
    }
    
    
    # Build df for plotting: variable of interest, then grouping vars.
    # Respecting the condition if present
    if (length(grouping_vars) == 0) {
      if (!is.null(cond_expr)) {
        df_for_plots <- data |>
          dplyr::filter(!!cond_expr) |>
          dplyr::select(dplyr::all_of(var))
      } else {
        df_for_plots <- data |>
          dplyr::select(dplyr::all_of(var))
      }
    } else {
      if (!is.null(cond_expr)) {
        df_for_plots <- data |>
          dplyr::filter(!!cond_expr) |>
          dplyr::select(dplyr::all_of(c(var, grouping_vars)))
      } else {
        df_for_plots <- data |>
          dplyr::select(dplyr::all_of(c(var, grouping_vars)))
      }
    }
    
    
    # Tweak df_for_plots to mask out-of-range numeric values to NA for plotting/summaries only
    if (v$type == "numeric" && !is.null(range_spec)) {
      df_for_plots <- df_for_plots |>
        dplyr::mutate(
          "{var}" := dplyr::if_else(
            !is.na(.data[[var]]) &
              (
                (.data[[var]] <  range_spec$lower) |
                  (.data[[var]] >  range_spec$upper) |
                  (!range_spec$lower_incl & .data[[var]] == range_spec$lower) |
                  (!range_spec$upper_incl & .data[[var]] == range_spec$upper)
              ),
            NA_real_,
            .data[[var]]
          )
        )
    }
    
    
    cat("\n\n## ",
        stringr::str_replace_all(label, "[\r\n]+", " ") |> stringr::str_squish(),
        "\n\n",
        sep = "")
    
    cat("\n\n **Validation info and details of non-allowed responses** \n\n")
    
    # validation summary and descriptive stats
    v |>
      dplyr::select(variable, type, colo,
                    allowed) |>
      kable_narrow() |>
      as.character() |>
      cat()

    v |>
      dplyr::select(`Allowed values specified?` = allowed_specified,
                    `Non-allowed (n)` = n_non_allowed,
                    `Missing (n)` = n_missing,
                    `Branching condition` = condition,
                    `Fail condition (n)` = n_fail_cond,
                    `Meet condition (n)` =  n_meet_cond,
                    `Meet condition & non-allowed (n)` = n_meet_cond_non_allowed,
                    `Meet condition & missing (n)` = n_meet_cond_missing) |>
      dplyr::mutate(dplyr::across(tidyselect::everything(), as.character)) |>
      tidyr::pivot_longer(
        everything(),
        names_to = "Metric",
        values_to = "Value"
      ) |>
      kable_narrow() |>
      as.character() |>
      cat()
    
    
    
    
    if(v$type == "factor" && !v$allowed_specified){
      cat(report_notice(
        "No allowed_values specified in the dictionary for this variable. Levels were inferred from the data."
      ))
    }
    
    
    # Print non-allowed values
    
    inv <- validation_log[[var]]$non_allowed_freq[[1]]
    inv_fail <- validation_log[[var]]$fail_cond_freq[[1]]
    
    #print(str(inv))
    
    
    
    # inv is always a list (possibly empty)
    if (!is.null(inv) && length(inv) > 0) {
      
      
      ## Factor invalids live in inv$values
      df_inv <- freq_to_df(inv$values, value_col = "Raw value")
      if (!is.null(df_inv)) {
        df_inv |>
          kable_narrow(caption = paste0("Non-allowed responses (condition satisfied)")) |>
          as.character() |>
          cat()
      }
      
      ## Numeric coercion failures live in inv$coercion
      df_inv <- freq_to_df(inv$coercion, value_col = "Raw value")
      if (!is.null(df_inv)) {
        df_inv |>
          kable_narrow(caption = paste0("Non-allowed numeric responses (could not be parsed)")) |>
          as.character() |>
          cat()
      }
      
      ## Numeric out-of-range values live in inv$out_of_range
      df_inv <- freq_to_df(inv$out_of_range, value_col = "Numeric value")
      if (!is.null(df_inv)) {
        df_inv |>
          kable_narrow(caption = paste0("Non-allowed numeric responses (out of allowed range)")) |>
          as.character() |>
          cat()
      }
      
      
    }
    
    
    
    # print responses where condition NOT satisfied
    if (!is.null(inv_fail) && length(inv_fail) > 0) {
      
      df_fail <- freq_to_df(inv_fail$values, value_col = "Raw value")
      
      if (!is.null(df_fail)) {
        df_fail |>
          kable_narrow(caption = "Responses where condition NOT satisfied") |>
          as.character() |>
          cat()
      }
    }
    
    
    
    # if (v$type == "numeric" && !is.null(range_spec) && v$n_non_allowed > 0) {
    #   cat(report_notice(
    #     "Numeric values outside the allowed range are included in validation ",
    #     "counts but excluded from summaries and plots.",
    #     level = "NOTE"
    #   ))
    # }
    
    
    
    
    if (!is.null(cond_expr)) {
      cat("\n\n **Summaries and plots below are restricted to allowed values where the condition is satisfied.** \n\n")
    }
    
    
    
    
    # a table of values for factors, otherwise a skim() of the data
    if (is.factor(df_for_plots[[var]])) {
      tab <- janitor::tabyl(df_for_plots[[var]]) |>
        janitor::adorn_totals() |>
        janitor::adorn_pct_formatting(digits = 1) |>
        as.data.frame() |>
        rename_NA_first_col("Missing / Non-allowed") |>
        (\(x) kable_narrow(x,
                           col.names = c("\u00A0", names(x)[-1]),
                           caption = "Distribution of responses"))() |>
        as.character() |>
        cat()
    } else {
      
      skimr::skim(df_for_plots[[var]]) |>
        kable_narrow() |>
        as.character() |>
        cat()
    }
    
    
    # if a text variable, just list some and be done with it (i.e. no plots etc)
    if ((dict |> dplyr::filter(variable_name == var) |> dplyr::pull(data_type)) == "text"){
      
      
      
      # Treat sentinel values as missing and drop them
      missing_codes_chr <- c("-999", "-888", "-777")
      
      df_for_plots <- df_for_plots |>
        dplyr::filter(!(.data[[var]] %in% missing_codes_chr))
      
      
      cat("Here are the responses and the grouping variables. There are ",
          nrow(df_for_plots),
          " responses here, excluding the -999 etc sentinel codes. \n\n")
      
      # set up and print an interactive table of responses
      col_width <- setNames(list(reactable::colDef(width = 400)), names(df_for_plots)[1])
      
      child <- paste0(
        "```{r echo=FALSE, progress=FALSE}\n",
        "reactable::reactable(df_for_plots, wrap=TRUE, columns = col_width, showPageSizeOptions = TRUE, pageSizeOptions = c(10,25,50,100), defaultPageSize = 10, fullWidth = FALSE, striped=TRUE, compact=TRUE, filterable=TRUE, resizable = TRUE)\n",
        "```\n"
      )
      out <- knitr::knit_child(text = child, envir = environment(), quiet=TRUE)
      cat(out, sep = "\n")
      
      next
    }
    
    # if a numerical variable, do some boxplotting
    #if (dict$data_type[dict$variable_name == var] == "numeric"){
    if ((dict |> dplyr::filter(variable_name == var) |> dplyr::pull(data_type)) == "numeric"){
      num_var <- names(df_for_plots)[1]
      group_vars <- names(df_for_plots)[-1]
      
      #first the overall plot
      df_for_plots$overall_group <- factor("Overall")
      p_overall <- OMESurvey::OME_boxplot_(
        data = df_for_plots,
        value_var = num_var,
        group_var = "overall_group",
        omitGroupLabels = TRUE,
        title = label
      )
      
      # ensure fig path exists
      fig_dir <- knitr::opts_chunk$get("fig.path")
      if (is.null(fig_dir) || fig_dir == "") fig_dir <- "figure/"
      if (!dir.exists(fig_dir)) dir.create(fig_dir, recursive = TRUE)
      
      # create a unique, stable filename (using section/variable/plot indices)
      fname <- file.path(fig_dir, sprintf("sec_%s_var_%02d_plot%02d.png",
                                          sec_file_key, i, 0))
      fname <- normalizePath(fname, winslash = "/", mustWork = FALSE)
      
      # save the plot at the computed size
      ggplot2::ggsave(fname, plot = p_overall,
                      width = 8, height = 3,
                      units = "in", dpi = 300)
      
      # emit markdown image tag (works inside results='asis')
      cat(sprintf("![](%s){width=100%%}\n\n", fname))
      
      # now for grouping variables
      for (j in seq_along(group_vars)) {
        g <- group_vars[j]
        # j is the numeric index, g is the name
        p <- OMESurvey::OME_boxplot_(
          data = df_for_plots,
          value_var = num_var,
          group_var = g,
          title = paste0(label, "\n by ", g)
        )
        
        
        # create a unique, stable filename (using section/variable/plot indices)
        fname <- file.path(fig_dir, sprintf("sec_%s_var_%02d_plot%02d.png",
                                            sec_file_key, i, j))
        fname <- normalizePath(fname, winslash = "/", mustWork = FALSE)
        
        # save the plot at the computed size
        ggplot2::ggsave(fname, plot = p,
                        width = 8, height = 1 + 0.3*nlevels(df_for_plots[[g]]),
                        units = "in", dpi = 300)
        
        # emit markdown image tag (works inside results='asis')
        cat(sprintf("![](%s){width=100%%}\n\n", fname))
        
      }
      
      next
    }
    
    
    colo_arg <- v$colo[[1]]
    
    # wrapper function for OME_stacked_bar() that just forwards all arguments; but wraps in a try/catch (with nice error-printing)
    OME_stacked_bar_tryCatch <- function(dat,
                                         response_var,
                                         group_var = NULL,
                                         facet_var = NULL,
                                         ...) {
      #print("WRAPPER DEBUG:")
      #print(response_var)
      #print(typeof(response_var))
      
      tryCatch({
        
        # Call the (programmatic _() version of the) function
        res <- OME_stacked_bar_(
          dat          = dat,
          response_var = response_var,
          group_var    = group_var,
          facet_var    = facet_var,
          ...
        )
        
        
        # if ggplot save and add to document
        if (inherits(res, "ggplot")) {
          fig_dir <- knitr::opts_chunk$get("fig.path")
          if (is.null(fig_dir) || fig_dir == "") fig_dir <- "figure/"
          if (!dir.exists(fig_dir)) dir.create(fig_dir, recursive = TRUE)
          fname <- file.path(
            fig_dir,
            sprintf("sec_%s_var_%02d_plot%02d.png",
                    stringr::str_extract(sec, "^[^:]+"), i, j)
          )
          fname <- normalizePath(fname, winslash = "/", mustWork = FALSE)
          
          ggplot2::ggsave(fname, plot = res,
                          width = 8, height = 3,
                          units = "in", dpi = 300)
          
          cat(sprintf("![](%s){width=100%%}\n\n", fname))
        } else{
          cat("Not a ggplot object")
          class(res)
        }
        invisible(res)
        
      }, # if there was an error, print it
      error = function(e) {
        #message("ERROR: ", conditionMessage(e))
        #message("TRACEBACK:")
        #print(sys.calls())
        #NULL
        
        cat(report_notice(
          "Error drawing variable plot: ", e$message
        ))
        invisible(NULL)
      })
    }
    
    # use this wrapper function with data and grouping vars as appropriate
    nvars <- ncol(df_for_plots)
    
    for (j in seq_len(nvars)) {
      sub_df <- df_for_plots[, unique(c(1, j)), drop = FALSE]
      #print(names(sub_df))
      
      plot_title <-
        paste0(
          label,
          if (j > 1) paste0("\n by ", names(sub_df)[2])
        )
      
      # Build symbols for tidy-eval
      resp_name <- names(sub_df)[1]
      grp_name  <- if (ncol(sub_df) > 1) names(sub_df)[2] else NULL
      
      OME_stacked_bar_tryCatch(
        dat          = sub_df,
        response_var = resp_name,
        group_var    = grp_name,
        percCut      = percCutVert,
        colo         = colo_arg,
        fillLabText  = NULL,
        propLabText  = NULL,
        titleText    = plot_title,
        NA_label     = "Missing / Non-allowed"
      )
    }
    
    
  } # end per-variable loop
} # end section loop
```
