---
title: "`r params$output_title`"
author: "`r params$output_author`"
date: "`r params$output_date`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
    toc_float: true
    code_folding: hide
    theme: cosmo
params:
  data_path: NULL
  dict_path: NULL
  dict_sheet: NULL
  est_chars_path: NULL
  est_chars_sheet: NULL
  est_char_vars: NULL
  est_char_types: NULL
  est_char_values: NULL
  est_char_statements: NULL
  output_title: NULL
  output_author: NULL
  output_date: NULL
  verbose: NULL
  percCutHoriz: NULL
  percCutVert: NULL
---

## General notes re this report

THINGS IN THIS SECTION MAY NEED REVISION/RELOCATION

This semi-automated survey summary/report contains the following:

- Any notes or warnings from data and dictionary importing, preparing, etc.
- A short summary of missingness.
- Sections of survey responses (grouped into sections by the preprocessor) which contain
  - A summary plot of all variables in the section, **if** they all have the same allowed values
  - Per-variable summaries containing (a) a summary of all data for the variable, (b) stats and a plot of values where any condition (for branching) is satisfied, (c) plots of values where any condition (for branching) is satisfied, broken down by any grouping variables present.
  - Numerical variables are plotted using boxplots and factors/categorical variables using stacked bar charts. Stacked bar charts show the percentage of response in each category, but these are suppressed if below `r params$percCutHoriz`% (section summaries) or `r params$percCutVert`% (variable-level plots).

  
  
The summary of all data include a 'response diagnostic table'. Here are some notes about it:

- 'NA (missing in data)' counts include only truly missing (NA) values, of which there should be none. (They should probably be -999 or -888).
- 'Non‑allowed' counts include values present but violating dictionary or type constraints. (If any of these are present they are listed in a separate frequency table.)
- The bold entries in the table are the ones that we expect might be non-zero.
- Any variables with non-zero entries in non-bold positions are listed at the very end of the document, in the section "Variables with unexpected values". It's a good idea to check that there aren't any there (or that any there are known about and are ok).

(In variables with a condition, there will often be both -999 and -888 responses. The -888s are those who answered the branching question and did not get branched to the present question. The -999s either (a) answered the branching question, were branched to this question and didn't answer it or (b) didn't answer the branching question.)

In the plots in this document the following approach is used with regard to dealing with missing data. (The intention is to have the plots focus on the distribution of the valid/allowed responses, but have info about missingness easily available.)

- Responses are only included if they are from a respondent who satisfied the branching condition.
- Only allowed values are plotted - non-allowed values as well as -999 (missing) and -777 (invalid) values indicated via the numbers in brackets.
- The numbers in brackets are of the form '(n/N)'. n is the number of allowed values plotted and N is the number of respondents shown the question (either everyone or everyone who satisfied the condition for branching to the question).

For example, suppose that in a survey with 100 respondents, a particular question has responses as follows: 25 Yes, 25 No, 25 don't answer(coded -999), 25 weren't branched to the question (coded -888). Then the plot will show 50% Yes, 50% No and have (50/75) at the side; to indicate that 50 allowed responses are plotted out of 75 respondents who were shown the question.


```{r results='asis'}
if (!requireNamespace("cowplot", quietly=TRUE)){
  cat(report_notice(
    "Package 'cowplot' is not detected. It is needed to centre legends underneath plots with respect to the whole figure rather than just the plot area. Any section summary plots with long labels on the vertical axis might benefit from  install.packages('cowplot') and then re-running.",
    level = "NOTE"))
}
```

## Notes/warnings from data & dictionary preparation, cleaning, etc

```{r setup, include=FALSE, echo=FALSE}
#knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.height=4)
knitr::opts_chunk$set(fig.height=2.5)



```




```{r}
# NEEDS to execute in global environment, not a clean one.
#cat("[1/4] Loading and checking inputs...")
#flush.console()

read_output <- survey_read_inputs(
  data_path = params$data_path,
  dict_path = params$dict_path,
  dict_sheet = params$dict_sheet,
  est_chars_path = params$est_chars_path,
  est_chars_sheet = params$est_chars_sheet,
  est_char_vars = params$est_char_vars,
  est_char_types = params$est_char_types,
  est_char_values = params$est_char_values,
  est_char_statements = params$est_char_statements
)


data <- read_output$data
dict <- read_output$dict
read_messages <- read_output$messages

#cat("[1/4] Done")
#flush.console()


#If verbose spit out the dictionary
if (params$verbose){
  print("VERBOSE MODE: here is the data dictionary")
  
  dict |>
    dplyr::mutate(dplyr::across(dplyr::where(is.character), # in character columns change newline (many possibilities) to double-space
                                ~ stringr::str_replace_all(., "\\r?\\n", "  "))) |>
    reactable::reactable(wrap=TRUE,
                         #columns = col_width,
                         showPageSizeOptions = TRUE,
                         pageSizeOptions = c(10,25,50,100),
                         defaultPageSize = 10,
                         fullWidth = FALSE,
                         striped=TRUE,
                         compact=TRUE,
                         filterable=TRUE,
                         resizable = TRUE)
  }

```



```{r results='asis'}
for (m in read_messages) {
  cat(report_notice(m$text, level = m$level))
}
```







```{r}

#cat("[2/4] Coercing and validating data...")
#flush.console()

prep_out <- survey_data_prepare(data, dict)

data <- prep_out$data
validation_log <- prep_out$validation_log
validation_df <- prep_out$validation_df
prep_messages <- prep_out$messages

#cat("[2/4] Done")
#flush.console()
```


```{r}
if (params$verbose){
  print("VERBOSE MODE: here is the validation log (in tibble, not list, form)")

  validation_df |>
    dplyr::mutate(dplyr::across(dplyr::where(is.character), # in character columns change newline (many possibilities) to double-space
                                ~ stringr::str_replace_all(., "\\\\r\\\\n|\\r\\n|\\n|\\r", "  "))) |>
    reactable::reactable(wrap=TRUE,
                         #columns = col_width,
                         showPageSizeOptions = TRUE,
                         pageSizeOptions = c(10,25,50,100),
                         defaultPageSize = 10,
                         fullWidth = FALSE,
                         striped=TRUE,
                         compact=TRUE,
                         filterable=TRUE,
                         resizable = TRUE)
}
```

```{r}  
if (params$verbose){
  print("VERBOSE MODE: here is the str() of the data")

  data |> str(strict.width="cut")
}
```

```{r results='asis'}
for (m in prep_messages) {
  cat(report_notice(m$text, level = m$level))
}
```



```{r}

percCutHoriz <- params$percCutHoriz
percCutVert <- params$percCutVert


#cat("[3/4] Making report...")
#flush.console()
```






# Missingness

A short summary of section-by-section complete missingness.

- In this section, missing means explicit non‑response codes (−999/−888). In the validation summaries, these are treated separately (as "non-allowed"). 
- A **section missing** means that **all questions in that section have no substantive response**.

This is a bit crude as it doesn't allow for branching; but the intention is to give a sense of survey attrition/dropout rates.

```{r}
# helper function
parse_report_sec <- function(sec) {
  s <- stringr::str_squish(as.character(sec))

  # Special case: Grouping variables (no prefix, always first)
  if (tolower(s) == "grouping variables") {
    return(list(
      display  = "Grouping variables",
      letter   = NA_character_,
      digit    = NA_integer_,
      code     = "GV",       # filename-safe short code
      sort_key = "0",        # always first
      file_key = "GV"
    ))
  }

  # General case:
  # Allow "A: Title" OR "A1: Title" where digit is 1-9
  m <- stringr::str_match(s, "^([A-Z])([1-9][0-9]?)?:\\s*(.+)$")
  if (!is.na(m[1, 1])) {
    letter <- m[1, 2]
    digit  <- if (!is.na(m[1, 3])) as.integer(m[1, 3]) else NA_integer_
    title  <- stringr::str_squish(m[1, 4])

    code <- if (is.na(digit)) letter else paste0(letter, digit)

    # sort: letter first, then digit (blank treated as 0, so "A:" comes before "A1:")
    digit_sort <- if (is.na(digit)) 0L else digit

    return(list(
      display  = paste0(code, ": ", title),
      letter   = letter,
      digit    = digit,
      code     = code,
      sort_key = sprintf("1_%s_%02d", letter, digit_sort),
      file_key = code
    ))
  }

  stop(
    "Invalid report_sec: '", s, "'. Allowed forms are:\n",
    "  - 'Grouping variables'\n",
    "  - 'X: Section title' where X is A–Z\n",
    "  - 'Xy: Section title' where X is A–Z and y is an integer 1–99 (no leading zero)"
  )
}
```


```{r}
# ------------------------------------------------------------------
# Section-level non-response summary
# A response is missing if it is NA OR one of the sentinel codes
# ------------------------------------------------------------------

missing_codes <- c(-999, -888)


# Map sections to variables (excluding Grouping variables),
# ordered in the same way as the main report
sec_meta_missing <-
  validation_df |>
  dplyr::filter(
    !is.na(report_sec),
    report_sec != "Grouping variables"
  ) |>
  dplyr::distinct(report_sec) |>
  dplyr::rename(sec_raw = report_sec) |>
  dplyr::rowwise() |>
  dplyr::mutate(
    info = list(parse_report_sec(sec_raw)),
    sec_sort_key = info$sort_key,
    sec_display  = info$display
  ) |>
  dplyr::ungroup() |>
  dplyr::arrange(sec_sort_key)

sec_vars <-
  validation_df |>
  dplyr::filter(
    !is.na(report_sec),
    report_sec != "Grouping variables"
  ) |>
  dplyr::left_join(
    sec_meta_missing |>
      dplyr::select(report_sec = sec_raw, sec_sort_key),
    by = "report_sec"
  ) |>
  dplyr::arrange(sec_sort_key, report_sec_ord) |>
  dplyr::summarise(
    vars = list(variable),
    .by = c(report_sec, sec_sort_key)
  ) |>
  dplyr::arrange(sec_sort_key) |>
  dplyr::select(-sec_sort_key)


data_miss_flags <- data
n_resp <- nrow(data)

# ---- Base section-level flags: all_missing_sec_i ----
for (i in seq_len(nrow(sec_vars))) {
  
  v_all <- sec_vars$vars[[i]]
  v_present <- v_all[v_all %in% names(data)]
  nm <- paste0("all_missing_sec_", i)
  
  if (length(v_present) == 0) {
    # Section exists conceptually but has no variables in data
    data_miss_flags[[nm]] <- rep(FALSE, n_resp)
    next
  }
  
  m <- as.matrix(
    as.data.frame(
      lapply(
        data[v_present],
        function(x) is.na(x) | (x %in% missing_codes)
      )
    )
  )
  
  data_miss_flags[[nm]] <- rowSums(!m) == 0
}

# ---- Cumulative flags: all_missing_sec_i_on (ALWAYS created) ----
sec_cols <- paste0("all_missing_sec_", seq_len(nrow(sec_vars)))

for (i in seq_along(sec_cols)) {
  data_miss_flags[[paste0(sec_cols[i], "_on")]] <-
    Reduce(`&`, data_miss_flags[sec_cols[i:length(sec_cols)]])
}



# ---- Aggregate summary ----

miss_summary <-
  data_miss_flags |>
  dplyr::summarise(
    dplyr::across(
      dplyr::matches("^all_missing_sec_\\d+$"),
      ~ sum(.x),
      .names = "{.col}"
    ),
    dplyr::across(
      dplyr::matches("^all_missing_sec_\\d+_on$"),
      ~ sum(.x),
      .names = "{.col}"
    )
  ) |>
  tidyr::pivot_longer(
    tidyselect::everything(),
    names_to = "key",
    values_to = "num"
  ) |>
  dplyr::mutate(
    type = dplyr::if_else(
      grepl("_on$", key),
      "This section and after",
      "This section only"
    ),
    sec_index = readr::parse_number(key),
    sec = factor(
      sec_vars$report_sec[sec_index],
      levels = sec_vars$report_sec
    ),
    prop = num / n_resp
  )

```

```{r}

# ---- Table ----


miss_table <-
  miss_summary |>
  dplyr::select(sec, type, num, prop) |>
  tidyr::pivot_wider(
    names_from = type,
    values_from = c(num, prop)
  )



miss_table |>
  dplyr::mutate(
    `This section only – Proportion`
      = formatC(`prop_This section only`, digits = 3, format = "f"),
    `This section and after – Proportion`
      = formatC(`prop_This section and after`, digits = 3, format = "f")
  ) |>
  dplyr::select(
    Section = sec,
    `This section – N` = `num_This section only`,
    `This section – Proportion` = `This section only – Proportion`,
    `This section and after – N` = `num_This section and after`,
    `This section and after – Proportion` = `This section and after – Proportion`
  ) |>
  dplyr::bind_rows(
    tibble::tibble(
      Section = "Total",
      `This section – N` = n_resp,
      `This section – Proportion` = "--",
      `This section and after – N` = n_resp,
      `This section and after – Proportion` = "--"
    )
  ) |>
  kable_narrow()
```

```{r fig.height=6, fig.width=8}

# ---- Plot ----


miss_summary |>
  ggplot2::ggplot(
    ggplot2::aes(
      x = sec,
      y = prop,
      colour = type,
      group = type
    )
  ) +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::labs(
    x = NULL,
    y = "Proportion of respondents",
    colour = NULL,
    title = "Section-level complete non-response"
  ) +
  OMESurvey::theme_OME() +
  ggplot2::theme(
    axis.text.x = ggplot2::element_text(angle = -30, vjust=1, hjust = 0),
    legend.position = "right",
    legend.direction = "vertical"
  ) +
  ggplot2::scale_colour_discrete(
    labels = c(
      "This section only" = "This section\nonly",
      "This section and after" = "This section\nand after"
    )
  )

```




```{r}
emit_reactable <- function(df, ...) {
  child <- paste0(
    "```{r echo=FALSE}\n",
    "reactable::reactable(df, ...)\n",
    "```\n"
  )
  out <- knitr::knit_child(text = child, envir = environment(), quiet = TRUE)
  cat(out, sep = "\n")
}



freq_to_df <- function(freq, value_col = "Value", add_total = TRUE) {
  if (is.null(freq) || length(freq) == 0) return(NULL)

  df <- data.frame(
    setNames(list(names(freq)), value_col),
    Frequency = as.integer(freq),
    check.names = FALSE
  )

  if (isTRUE(add_total)) {
    df <- rbind(
      df,
      data.frame(
        setNames(list("Total"), value_col),
        Frequency = sum(df$Frequency, na.rm = TRUE),
        check.names = FALSE
      )
    )
  }

  df
}

remove_common_prefix <- function(x) {
  # find longest common prefix length
  pref_len <- 0L
  for (len in seq_len(min(nchar(x)))) {
    if (all(substr(x, 1, len) == substr(x[1], 1, len))) {
      pref_len <- len
    } else break
  }
  # build results
  if (pref_len == 0L) {
    list(prefix = "", remainder = x)
  } else {
    prefix <- substr(x[1], 1, pref_len)
    remainder <- substring(x, pref_len + 1)
    list(prefix = prefix, remainder = remainder)
  }
}






```



```{r results='asis', echo=FALSE, message=TRUE, fig.width=10, fig.height=6}

# order sections and ensure variables are ordered within each section

sections_raw <-
  validation_df |>
  dplyr::pull(report_sec) |>
  as.character() |>
  stats::na.omit() |>
  unique()

sec_meta <-
  tibble::tibble(sec_raw = sections_raw) |>
  dplyr::rowwise() |>
  dplyr::mutate(info = list(parse_report_sec(sec_raw))) |>
  dplyr::mutate(
    sec_display  = info$display,
    sec_letter   = info$letter,
    sec_sort_key = info$sort_key,
    sec_file_key = info$file_key
  ) |>
  dplyr::ungroup()




# Validate: section codes must be unique
# (e.g. "A", "A1", "B2", "GV" must not repeat)

dups <- sec_meta |>
  dplyr::group_by(sec_file_key) |>
  dplyr::filter(dplyr::n() > 1) |>
  dplyr::summarise(
    sections = paste(sprintf('"%s"', sec_display), collapse = ", "),
    .groups = "drop"
  )

if (nrow(dups) > 0) {
  msg <- paste(
    sprintf(
      'Key "%s" is used by sections: %s',
      dups$sec_file_key,
      dups$sections
    ),
    collapse = "\n"
  )
  
  stop(
    "Duplicate section codes (file keys) found:\n\n",
    msg,
    "\n\nEach code (before the colon in the section title) must be unique."
  )
}


# Order: Grouping variables first (sort_key "0"), then A, B, C...
sec_meta <- sec_meta |>
  dplyr::arrange(sec_sort_key)

sections <- sec_meta$sec_display




vars_with_unexpected_nonzero <- list()

for (sec in sections){
  cat("\n\n# ", sec, "\n\n", sep = "")
  
  # catch potential file-naming issues
  sec_file_key <- sec_meta$sec_file_key[match(sec, sec_meta$sec_display)]
  if (is.na(sec_file_key)) stop("Could not determine file key for section: ", sec)
  
  
  vars_in_sec <-
    validation_df |>
    dplyr::filter(report_sec == sec) |>
    dplyr::arrange(report_sec_ord)
  
  vars <- vars_in_sec$variable
  #vars |> print()
  
  if (length(vars) == 0) next
  
  # vars_in_sec |>
  #   dplyr::select(type, allowed_specified, allowed_display) |> print()
  # 
  # vars_in_sec |>
  #   dplyr::select(type, allowed_specified, allowed_display) |>
  #   dplyr::distinct() |> print()
  
  
  #if not relevant, make a note that there's no summary plot, otherwise do it
  if (length(vars) == 1 || (vars_in_sec |>
                            dplyr::select(type, allowed_specified, allowed_display) |>
                            dplyr::distinct() |>
                            nrow()) != 1
  ) {
    cat(report_notice(
      "No section summary plot for this section.",
      level = "NOTE"
    ))
  } else {
    
        # determine section type
    sec_type <- vars_in_sec$type[[1]]
    #sec_type |> print()

    
    # cat(report_notice(
    #   "For numeric variables, values outside the allowed range are excluded from plots. ",
    #   "For categorical variables, non-allowed values are implicitly excluded via factor coercion. ",
    #   "Validation tables include these values.",
    #   level = "NOTE"
    # ))
    
    # prepare df and labels for plotting

raw_vars <- paste0("raw_", vars)

df_plot <- data |> dplyr::select(dplyr::any_of(c(vars, raw_vars)))
    
    # Apply numeric range masking for section-level plots
    for (k in seq_len(nrow(vars_in_sec))) {
      
      vname <- vars[k]
      raw_name <- paste0("raw_", vname)
      
      allowed_vals <- vars_in_sec$allowed_technical[[k]]
      cond_k <- vars_in_sec$condition[k]
      
      

# ALWAYS use data here
raw_vec   <- data[[raw_name]]
value_vec <- data[[vname]]

# include mask, using raw variables to evaluate the condition

include_vec <- if (!is.na(cond_k) && nzchar(cond_k)) {
  
  cond_k_expr <- rlang::parse_expr(cond_k)
  cond_k_expr_raw <- OMESurvey::rewrite_cond_to_raw(cond_k_expr, names(data))
  
  tmp <- data |>
    dplyr::mutate(.tmp = !!cond_k_expr_raw) |>
    dplyr::pull(.tmp)
  
  if (anyNA(tmp)) {
    cond_k_expr_txt     <- rlang::expr_text(cond_k_expr)
    cond_k_expr_raw_txt <- rlang::expr_text(cond_k_expr_raw)
    
    cat(report_notice(
      paste0(
        "Section-summary condition for variable '", vname,
        "' evaluated to NA for ", sum(is.na(tmp)),
        " row", if (sum(is.na(tmp)) == 1) "" else "s",
        ". These rows are being treated as condition not met.\n",
        "Original condition: ", cond_k_expr_txt, "\n",
        "Raw-evaluated condition: ", cond_k_expr_raw_txt
      ),
      level = "WARNING"
    ))
    
    tmp[is.na(tmp)] <- FALSE
  }
  
  tmp
} else {
  rep(TRUE, nrow(data))
}


# exclude sentinels
include_vec <- include_vec & !(raw_vec %in% c("-888", -888))

# plot mask
has_allowed <- !is.null(allowed_vals) && length(allowed_vals) > 0

if (sec_type == "factor") {
  plot_vec <- if (has_allowed) {
    raw_vec %in% allowed_vals
  } else {
    !is.na(value_vec)
  }
} else {
  plot_vec <- !is.na(value_vec)
}


# attach
df_plot[[paste0(vname, "_value")]]   <- df_plot[[vname]]
df_plot[[paste0(vname, "_include")]] <- include_vec
df_plot[[paste0(vname, "_plot")]]    <- plot_vec

      
    }
    
    # remove original variables (now _value)
    df_plot <- df_plot |> dplyr::select(-all_of(vars))
    
df_plot <- df_plot |>
  dplyr::select(-starts_with("raw_"))


    labels_vec <-
      as.character(vars_in_sec$item_statement) |>
      setNames(vars_in_sec$variable)
    
    
    # get common prefix out of labels_vec to use as title
    labels_prefix <- remove_common_prefix(labels_vec)
    title_text <- labels_prefix$prefix
    labels_vec <- labels_prefix$remainder
    
    
    
    # compute dynamic height: 2 inches + 0.5 inch per variable
    fig_h <- (2 + 0.5 * length(vars))
    
    
    
    
    
    tryCatch({
      
      if (sec_type == "factor") {
        
        # use colo and ordering from the first variable
        colo <- (vars_in_sec |> dplyr::pull(colo) |> utils::head(1))[[1]]
        order_values <- (vars_in_sec |> dplyr::pull(order_values) |> utils::head(1))[[1]]
        
        # factor section summary plot
        n_cat <- nlevels(df_plot[[1]])
        fill_label_width <-
          if (n_cat <= 5) 20 else max(8, floor(100 / n_cat))
        
        res <- OMESurvey::summary_plot_stacked_bar(
          df_plot,
          labels_vec,
          percCut = percCutHoriz,
          colo = colo,
          order_values = order_values,
          titleText = title_text,
          NA_label = "Missing / Non-allowed",
          fill_label_width = fill_label_width
        ) |>
          OMESurvey::centre_legend_below(rel_heights = c(1, 0.8 / fig_h))
        
      } else if (sec_type == "numeric") {
        
        # numeric section summary plot
        res <- OMESurvey::summary_plot_boxplot(
          df_plot,
          labels_vec = labels_vec,
          titleText = title_text
        )
        
      } else {
        res <- NULL
      }
      
      
      if (inherits(res, "ggplot")) {
        # ensure fig path exists
        fig_dir <- knitr::opts_chunk$get("fig.path")
        if (is.null(fig_dir) || fig_dir == "") fig_dir <- "figure/"
        if (!dir.exists(fig_dir)) dir.create(fig_dir, recursive = TRUE)
        
        # create a unique, stable filename (using section name)
        fname <- file.path(
          fig_dir,
          sprintf("sec_%s_plot_overall.png", sec_file_key)
        )
        fname <- normalizePath(fname, winslash = "/", mustWork = FALSE)
        
        # save the plot and emit markdown image tag (for inside an as-is chunk)
        ggplot2::ggsave(fname, plot = res,
                        width = 8, height = fig_h,
                        units = "in", dpi = 300)
        
        cat(sprintf("![](%s){width=100%%}\n\n", fname))
      }
      
      invisible(res)
      
    }, error = function(e) {
      cat(report_notice(
        "Error drawing section plot: ", e$message
      ))
      invisible(NULL)
    })
  }
  
  
  
  
  # now emit per-variable summaries for this section
  for (i in seq_len(nrow(vars_in_sec))){
    
    has_unexpected_nonzero <- FALSE
    
    v <- vars_in_sec[i, ] # the current row of vars_in_sec, in turn from validation_df. So has
    var <- v$variable
    label <- ifelse(is.na(v$item_statement) || v$item_statement == "", var, v$item_statement)
    
    
    # select grouping variables for this report; excluding the variable itself if appropriate
    grouping_vars <-
      dict |>
      dplyr::filter(grouping_var == TRUE) |>
      dplyr::pull(variable_name) |>
      setdiff(var)   # ensure the target var is not duplicated
    
    # ensure grouping_vars exist in data (safe select)
    grouping_vars <- grouping_vars[grouping_vars %in% names(data)]
    
    
    # extract condition for this variable
    cond <-
      dict |>
      dplyr::filter(variable_name == vars[i]) |>
      dplyr::pull(condition)
    
    # parse condition safely with tryCatch and friendly error message
    cond_expr <- NULL
    if (!is.na(cond) && nzchar(cond)) {
      cond_expr <- tryCatch(
        rlang::parse_expr(cond),
        error = function(e) {
          cat(report_notice(
            "Error parsing filter condition: ", e$message
          ))
          NULL
        }
      )
    }
    
    
    # Build df for plotting: variable of interest, then grouping vars.
    # Respecting the condition if present
    
    
    raw_name <- paste0("raw_", var)
    
    
#  build df (NO filtering yet) ----
if (length(grouping_vars) == 0) {
  df_for_plots <- data |>
    dplyr::select(dplyr::all_of(c(var, raw_name)))
} else {
  df_for_plots <- data |>
    dplyr::select(dplyr::all_of(c(var, raw_name, grouping_vars)))
}

    
#  include mask (instead of filter) ----
if (!is.null(cond_expr)) {
  
   cond_expr_raw <- rewrite_cond_to_raw(cond_expr, names(data))
   
   
  include_vec <- data |>
    dplyr::mutate(.tmp = !!cond_expr_raw) |>
    dplyr::pull(.tmp)
  
  n_na_cond <- sum(is.na(include_vec))


  if (n_na_cond > 0) {
    
    # Try to identify variables referenced in the condition
    cond_vars <- tryCatch(
      all.vars(cond_expr),
      error = function(e) character(0)
    )
    
    # Keep only names that are actually columns in the data
    cond_vars <- intersect(cond_vars, names(data))
    
    # Summarise missingness in those variables
    if (length(cond_vars) > 0) {
      cond_var_info <- vapply(
        cond_vars,
        function(vv) {
          x <- data[[vv]]
          
          # count NA
          n_na <- sum(is.na(x))
          
          # optionally also count sentinel values if relevant
          n_sentinel <- sum(x %in% c(-999, -888, -777, "-999", "-888", "-777"), na.rm = TRUE)
          
          paste0(
            vv, ": ",
            n_na, " NA",
            if (n_sentinel > 0) paste0(", ", n_sentinel, " sentinel") else ""
          )
        },
        FUN.VALUE = character(1)
      )
      
      cond_var_info_txt <- paste(cond_var_info, collapse = "; ")
      
      cat(report_notice(
        paste0(
          "Branching condition for variable '", var,
          "' evaluated to NA for ", n_na_cond, " row",
          if (n_na_cond == 1) "" else "s", ". ",
          "These rows are being treated as condition not met. ",
          "Condition: ", cond, ". ",
          "Variables referenced: ", paste(cond_vars, collapse = ", "), ". ",
          "Missingness in referenced variables: ", cond_var_info_txt, "."
        ),
        level = "WARNING"
      ))
      
    } else {
      
      # fallback if no variables could be extracted
      cat(report_notice(
        paste0(
          "Branching condition for variable '", var,
          "' evaluated to NA for ", n_na_cond, " row",
          if (n_na_cond == 1) "" else "s", ". ",
          "These rows are being treated as condition not met. ",
          "Condition: ", cond, "."
        ),
        level = "WARNING"
      ))
    }
  }


  include_vec[is.na(include_vec)] <- FALSE

} else {
  include_vec <- rep(TRUE, nrow(df_for_plots))
}

    
    
# plot mask (FULL DATA) ----
allowed_vals <- v$allowed_technical[[1]]
has_allowed <- !is.null(allowed_vals) && length(allowed_vals) > 0

if (v$type == "factor") {

  if (has_allowed) {
    plot_vec <- df_for_plots[[raw_name]] %in% allowed_vals
  } else {
    plot_vec <- !is.na(df_for_plots[[var]])
  }

} else if (v$type == "numeric" && has_allowed) {

  range_spec <- allowed_vals

  plot_vec <- !is.na(df_for_plots[[var]]) &
    (
      (df_for_plots[[var]] > range_spec$lower |
         (range_spec$lower_incl & df_for_plots[[var]] == range_spec$lower)) &
      (df_for_plots[[var]] < range_spec$upper |
         (range_spec$upper_incl & df_for_plots[[var]] == range_spec$upper))
    )

} else {

  plot_vec <- !is.na(df_for_plots[[var]])

}

# apply masking ----
df_for_plots[[var]][!plot_vec] <- NA

# NOW filter using include ----
df_for_plots <- df_for_plots[include_vec, , drop = FALSE]




    
    # if (length(grouping_vars) == 0) {
    #   if (!is.null(cond_expr)) {
    #     df_for_plots <- data |>
    #       dplyr::filter(!!cond_expr) |>
    #       dplyr::select(dplyr::all_of(c(var, raw_name)))
    #   } else {
    #     df_for_plots <- data |>
    #       dplyr::select(dplyr::all_of(c(var, raw_name)))
    #   }
    # } else {
    #   if (!is.null(cond_expr)) {
    #     df_for_plots <- data |>
    #       dplyr::filter(!!cond_expr) |>
    #       dplyr::select(dplyr::all_of(c(var, raw_name, grouping_vars)))
    #   } else {
    #     df_for_plots <- data |>
    #       dplyr::select(dplyr::all_of(c(var, raw_name, grouping_vars)))
    #   }
    # }
    # 
    # 
    # 
    # 
    # raw_name <- paste0("raw_", var)
    # 
    # if (v$type == "factor") {
    #   plot_vec <- df_for_plots[[raw_name]] %in% v$allowed_technical[[1]]
    # } else if (v$type == "numeric" && !is.null(v$allowed_technical[[1]])) {
    #   
    #   range_spec <- v$allowed_technical[[1]]
    #   
    #   plot_vec <- !is.na(df_for_plots[[var]]) &
    #     (
    #       (df_for_plots[[var]] >  range_spec$lower | 
    #          (range_spec$lower_incl & df_for_plots[[var]] == range_spec$lower)) &
    #         (df_for_plots[[var]] <  range_spec$upper | 
    #            (range_spec$upper_incl & df_for_plots[[var]] == range_spec$upper))
    #     )
    #   
    # } else {
    #   
    #   plot_vec <- !is.na(df_for_plots[[var]])
    #   
    # }
    # 
    # df_for_plots[[var]][!plot_vec] <- NA
    
    
    
    
    # # Tweak df_for_plots to mask out-of-range numeric values to NA for plotting/summaries only
    # range_spec <- v$allowed_technical[[1]]
    # 
    # if (v$type == "numeric" && !is.null(range_spec)) {
    #   df_for_plots <- df_for_plots |>
    #     dplyr::mutate(
    #       "{var}" := dplyr::if_else(
    #         !is.na(.data[[var]]) &
    #           (
    #             (.data[[var]] <  range_spec$lower) |
    #               (.data[[var]] >  range_spec$upper) |
    #               (!range_spec$lower_incl & .data[[var]] == range_spec$lower) |
    #               (!range_spec$upper_incl & .data[[var]] == range_spec$upper)
    #           ),
    #         NA_real_,
    #         .data[[var]]
    #       )
    #     )
    # }
    # 
    
    cat("\n\n## ",
        stringr::str_replace_all(label, "[\r\n]+", " ") |> stringr::str_squish(),
        "\n\n",
        sep = "")
    
    cat("\n\n Validation info \n\n")
    
    # validation summary and descriptive stats
    v |>
      dplyr::select(variable, type, #colo,
                    `Allowed values specified?` = allowed_specified,
                    `Allowed values` = allowed_display,
                    `Branching condition` = condition) |>
      kable_narrow() |>
      as.character() |>
      cat()

    # cat("\n\n table that can now be deleted? \n\n")
    # v |>
    #   dplyr::select(`Allowed values specified?` = allowed_specified,
    #                 `Non-allowed (n)` = n_non_allowed,
    #                 `Missing (n)` = n_missing,
    #                 `Branching condition` = condition,
    #                 `Fail condition (n)` = n_fail_cond,
    #                 `Meet condition (n)` =  n_meet_cond,
    #                 `Meet condition & non-allowed (n)` = n_meet_cond_non_allowed,
    #                 `Meet condition & missing (n)` = n_meet_cond_missing) |>
    #   dplyr::mutate(dplyr::across(tidyselect::everything(), as.character)) |>
    #   tidyr::pivot_longer(
    #     everything(),
    #     names_to = "Metric",
    #     values_to = "Value"
    #   ) |>
    #   kable_narrow() |>
    #   as.character() |>
    #   cat()
    
    has_valid_cond <- !is.na(v$condition)
    
    
    should_bold <- function(row, col, val, has_valid_cond) {
      # condition-met and 1-column rules
      if ((col == "Condition met" || (!has_valid_cond && col == "Total")) &&
          row %in% c("Allowed", "-999 (Missing)", "-777 (Invalid)")) return(TRUE)
      
      # condition-not-met rules
      if (col == "Condition not met" &&
          row %in% c("-999 (Missing)", "-888 (Not branched)")) return(TRUE)
      
      return(FALSE)
    }
    
    
      
    diag_tbl <- validation_log[[var]]$diag_table[[1]]
    diag_tbl_fmt <- diag_tbl
    
    # convert to character for safe formatting
    diag_tbl_fmt[] <- lapply(diag_tbl_fmt, as.character)
    
    # get grand total (bottom-right cell)
    grand_total <- suppressWarnings(
      as.numeric(diag_tbl[nrow(diag_tbl), "Total"])
    )
    
    
    # identify columns that have any boldable cells
    cols_with_boldable <- setNames(
      logical(ncol(diag_tbl)),
      colnames(diag_tbl)
    )
    
    for (col_idx in seq_len(ncol(diag_tbl))) {
      col_name <- colnames(diag_tbl)[col_idx]
      
      for (row_idx in seq_len(nrow(diag_tbl))) {
        row_name <- rownames(diag_tbl)[row_idx]
        
        if (should_bold(row_name, col_name, val=NA, has_valid_cond)) {
          cols_with_boldable[col_name] <- TRUE
          break
        }
      }
    }
    
    
    for (row_idx in seq_len(nrow(diag_tbl))) {
      for (col_idx in seq_len(ncol(diag_tbl))) {
        
        row_name <- rownames(diag_tbl)[row_idx]
        col_name <- colnames(diag_tbl)[col_idx]
        val <- as.numeric(diag_tbl[row_idx, col_idx])
        
        
        
        # detect unexpected non-zero within relevant columns
        if (cols_with_boldable[col_name] &&
            row_name != "Total" &&
            !should_bold(row_name, col_name, val, has_valid_cond) &&
            !is.na(val) && val > 0) {
          
          has_unexpected_nonzero <- TRUE
        }
        
        
        # add % to Total row & column
        
        if (!is.na(val) && grand_total > 0) {
          
          if (xor(row_name == "Total", col_name == "Total")) {
            pct <- round(100 * val / grand_total, 0)
            val_str <- paste0(val, " (", pct, "%)")
          } else {
            val_str <- as.character(val)
          }
          
        } else {
          val_str <- as.character(val)
        }
        
        
        # ---- Apply bold rule
        if (should_bold(row_name, col_name, val, has_valid_cond)) {
          diag_tbl_fmt[row_idx, col_idx] <- paste0("**", val_str, "**")
        } else {
          diag_tbl_fmt[row_idx, col_idx] <- val_str
        }
      }
    }
    
    
    
    cat("\n\nResponse diagnostic table (counts)\n\n")
    
    diag_tbl_fmt |>
      tibble::rownames_to_column("Category") |>
      kable_narrow() |>
      as.character() |>
      cat()

    
        
    if (v$type == "factor" && !v$allowed_specified){
      cat(report_notice(
        "No allowed_values specified in the dictionary for this variable. Levels were inferred from the data."
      ))
    }
    
    
    # Print non-allowed values
    
    inv <- validation_log[[var]]$non_allowed_freq[[1]]
    inv_fail <- validation_log[[var]]$fail_cond_freq[[1]]
    
    #print(str(inv))
    
    
    
    # inv is always a list (possibly empty)
    if (!is.null(inv) && length(inv) > 0) {
      
      
      ## Factor invalids live in inv$values
      df_inv <- freq_to_df(inv$values, value_col = "Raw value")
      
      if (!is.null(df_inv)) {
        df_inv <- df_inv[!(df_inv$`Raw value` %in% c("-999", "-888", "-777", "Total")), ]
        
        if (nrow(df_inv) > 0) {
          df_inv |>
            kable_narrow(caption = paste0("Non-allowed responses (condition satisfied)")) |>
            as.character() |>
            cat()
          
        }
      }
      
      
      ## Numeric coercion failures live in inv$coercion
      df_inv <- freq_to_df(inv$coercion, value_col = "Raw value")
      
      if (!is.null(df_inv)) {
        df_inv <- df_inv[!(df_inv$`Raw value` %in% c("-999", "-888", "-777", "Total")), ]
        
        if (nrow(df_inv) > 0) {
          df_inv |>
            kable_narrow(caption = paste0("Non-allowed numeric responses (could not be parsed)")) |>
            as.character() |>
            cat()
        }
      }
      
      ## Numeric out-of-range values live in inv$out_of_range
      df_inv <- freq_to_df(inv$out_of_range, value_col = "Raw value")
      
      # if something there then exclude the sentinel values
      if (!is.null(df_inv)) {
        df_inv <- df_inv[!(df_inv$`Raw value` %in% c("-999", "-888", "-777", "Total")), ]
        
        #if something still there then print it
        if (nrow(df_inv) > 0) {
        df_inv |>
          kable_narrow(caption = paste0("Non-allowed numeric responses (out of allowed range)")) |>
          as.character() |>
          cat()
        }
      }
      
      
    }
    
    
    
    # print responses where condition NOT satisfied
    if (!is.null(inv_fail) && length(inv_fail) > 0) {
      
      
      df_fail <- freq_to_df(inv_fail$values, value_col = "Raw value")
      
      # if something there then exclude the sentinel values
      if (!is.null(df_fail)) {
        df_fail <- df_fail[!(df_fail$`Raw value` %in% c("-999", "-888", "-777", "Total")), ]
        
        #if something still there then print it
        if (nrow(df_fail) > 0) {
          df_fail |>
            kable_narrow(caption = "Responses where condition NOT satisfied") |>
            as.character() |>
            cat()
          
        }
        
      }
    }
    
    
    
    # if (v$type == "numeric" && !is.null(range_spec) && v$n_non_allowed > 0) {
    #   cat(report_notice(
    #     "Numeric values outside the allowed range are included in validation ",
    #     "counts but excluded from summaries and plots.",
    #     level = "NOTE"
    #   ))
    # }
    
    
    
    
    if (!is.null(cond_expr)) {
      cat("\n\n **Summaries and plots below are restricted to allowed values where the condition is satisfied.** \n\n")
    }
    
    
    
    
    # a table of values for factors, otherwise a skim() of the data
    if (is.factor(df_for_plots[[var]])) {
      tab <- janitor::tabyl(df_for_plots[[var]]) |>
        janitor::adorn_totals() |>
        janitor::adorn_pct_formatting(digits = 1) |>
        as.data.frame() |>
        rename_NA_first_col("Missing / Non-allowed") |>
        (\(x) kable_narrow(x,
                           col.names = c("\u00A0", names(x)[-1]),
                           caption = "Distribution of responses"))() |>
        as.character() |>
        cat()
    } else {
      
      skimr::skim(df_for_plots[[var]]) |>
        kable_narrow() |>
        as.character() |>
        cat()
    }
    
    # stop(
    #   paste0(
    #     names(df_for_plots)
    #   )
    # )

    df_for_plots <- df_for_plots |>
  dplyr::select(-starts_with("raw_"))

    
    
    # if a text variable, just list and be done with it (i.e. no plots etc)
    if ((dict |> dplyr::filter(variable_name == var) |> dplyr::pull(data_type)) == "text"){
      
      
      
      # Treat sentinel values as missing and drop them
      missing_codes_chr <- c("-999", "-888", "-777")
      
      df_for_plots <- df_for_plots |>
        dplyr::filter(!(.data[[var]] %in% missing_codes_chr))
      
      
      cat("Here are the responses and the grouping variables. There are ",
          nrow(df_for_plots),
          " responses here, excluding the -999 etc sentinel codes. \n\n")
      
      # set up and print an interactive table of responses
      col_width <- setNames(list(reactable::colDef(width = 400)), names(df_for_plots)[1])
      
      child <- paste0(
        "```{r echo=FALSE, progress=FALSE}\n",
        "reactable::reactable(df_for_plots, wrap=TRUE, columns = col_width, showPageSizeOptions = TRUE, pageSizeOptions = c(10,25,50,100), defaultPageSize = 10, fullWidth = FALSE, striped=TRUE, compact=TRUE, filterable=TRUE, resizable = TRUE)\n",
        "```\n"
      )
      out <- knitr::knit_child(text = child, envir = environment(), quiet=TRUE)
      cat(out, sep = "\n")
      
      next
    }
    
    # if a numerical variable, do some boxplotting
    #if (dict$data_type[dict$variable_name == var] == "numeric"){
    if ((dict |> dplyr::filter(variable_name == var) |> dplyr::pull(data_type)) == "numeric"){
      num_var <- names(df_for_plots)[1]
      group_vars <- names(df_for_plots)[-1]
      
      #first the overall plot
      df_for_plots$overall_group <- factor("Overall")
      p_overall <- OMESurvey::OME_boxplot_(
        data = df_for_plots,
        value_var = num_var,
        group_var = "overall_group",
        omitGroupLabels = TRUE,
        titleText = label
      )
      
      # ensure fig path exists
      fig_dir <- knitr::opts_chunk$get("fig.path")
      if (is.null(fig_dir) || fig_dir == "") fig_dir <- "figure/"
      if (!dir.exists(fig_dir)) dir.create(fig_dir, recursive = TRUE)
      
      # create a unique, stable filename (using section/variable/plot indices)
      fname <- file.path(fig_dir, sprintf("sec_%s_var_%02d_plot%02d.png",
                                          sec_file_key, i, 0))
      fname <- normalizePath(fname, winslash = "/", mustWork = FALSE)
      
      # save the plot at the computed size
      ggplot2::ggsave(fname, plot = p_overall,
                      width = 8, height = 3,
                      units = "in", dpi = 300)
      
      # emit markdown image tag (works inside results='asis')
      cat(sprintf("![](%s){width=100%%}\n\n", fname))
      
      # now for grouping variables
      for (j in seq_along(group_vars)) {
        g <- group_vars[j]
        # j is the numeric index, g is the name
        p <- OMESurvey::OME_boxplot_(
          data = df_for_plots,
          value_var = num_var,
          group_var = g,
          titleText = paste0(label, "\n by ", g)
        )
        
        
        # create a unique, stable filename (using section/variable/plot indices)
        fname <- file.path(fig_dir, sprintf("sec_%s_var_%02d_plot%02d.png",
                                            sec_file_key, i, j))
        fname <- normalizePath(fname, winslash = "/", mustWork = FALSE)
        
        # save the plot at the computed size
        ggplot2::ggsave(fname, plot = p,
                        width = 8, height = 1 + 0.3*nlevels(df_for_plots[[g]]),
                        units = "in", dpi = 300)
        
        # emit markdown image tag (works inside results='asis')
        cat(sprintf("![](%s){width=100%%}\n\n", fname))
        
      }
      
      next
    }
    
    
    colo_arg <- v$colo[[1]]
    
    # wrapper function for OME_stacked_bar() that just forwards all arguments; but wraps in a try/catch (with nice error-printing)
    OME_stacked_bar_tryCatch <- function(dat,
                                         response_var,
                                         group_var = NULL,
                                         facet_var = NULL,
                                         sec_key,
                                         qn_idx,
                                         group_idx,
                                         ...) {
      tryCatch({
        
        # Call the (programmatic _() version of the) function
        res <- OMESurvey::OME_stacked_bar_(
          dat          = dat,
          response_var = response_var,
          group_var    = group_var,
          facet_var    = facet_var,
          ...
        )
        
        
        # if ggplot save and add to document
        if (inherits(res, "ggplot")) {
          fig_dir <- knitr::opts_chunk$get("fig.path")
          if (is.null(fig_dir) || fig_dir == "") fig_dir <- "figure/"
          if (!dir.exists(fig_dir)) dir.create(fig_dir, recursive = TRUE)
          fname <- file.path(
            fig_dir,
            sprintf("sec_%s_var_%02d_plot%02d.png",
                    sec_key, qn_idx, group_idx)
          )
          fname <- normalizePath(fname, winslash = "/", mustWork = FALSE)
          
          ggplot2::ggsave(fname, plot = res,
                          width = 8, height = 3,
                          units = "in", dpi = 300)
          
          cat(sprintf("![](%s){width=100%%}\n\n", fname))
        } else{
          cat("Not a ggplot object")
          class(res)
        }
        invisible(res)
        
      }, # if there was an error, print it
      error = function(e) {
        #message("ERROR: ", conditionMessage(e))
        #message("TRACEBACK:")
        #print(sys.calls())
        #NULL
        
        cat(report_notice(
          "Error drawing variable plot: ", e$message
        ))
        invisible(NULL)
      })
    }
    
    # use this wrapper function with data and grouping vars as appropriate
    nvars <- ncol(df_for_plots)
    
    for (j in seq_len(nvars)) {
      sub_df <- df_for_plots[, unique(c(1, j)), drop = FALSE]
      #print(names(sub_df))
      
      plot_title <-
        paste0(
          label,
          if (j > 1) paste0("\n by ", names(sub_df)[2])
        )
      
      # Build symbols for tidy-eval
      resp_name <- names(sub_df)[1]
      grp_name  <- if (ncol(sub_df) > 1) names(sub_df)[2] else NULL
      
      OME_stacked_bar_tryCatch(
        dat          = sub_df,
        response_var = resp_name,
        group_var    = grp_name,
                                         sec_key = sec_file_key,
                                         qn_idx = i,
                                         group_idx = j,
        na.rm        = TRUE,
        count_style  = "both",
        percCut      = percCutVert,
        colo         = colo_arg,
        fillLabText  = NULL,
        propLabText  = NULL,
        titleText    = plot_title,
        NA_label     = "Missing / Non-allowed"
      )
    }
    
    
    
    if (has_unexpected_nonzero) {
      if (is.null(vars_with_unexpected_nonzero[[sec]])) {
        vars_with_unexpected_nonzero[[sec]] <- var
      } else {
        vars_with_unexpected_nonzero[[sec]] <- c(vars_with_unexpected_nonzero[[sec]], var)
      }
    }
    
    
  } # end per-variable loop
} # end section loop



cat("\n\n# Variables with unexpected values \n\n")

if (length(vars_with_unexpected_nonzero) == 0) {
  cat("None.\n")
} else {
  for (sec_name in names(vars_with_unexpected_nonzero)) {
    
    vars_in_sec_flagged <- vars_with_unexpected_nonzero[[sec_name]]
    
    if (length(vars_in_sec_flagged) > 0) {
      cat("**", sec_name, "**", "\n", sep = "")
      cat(paste0("  ", vars_in_sec_flagged, collapse = "\n"), "\n\n")
    }
  }
}

```
