---
title: "Dave's plotting (and other) functions"
output:
  rmarkdown::html_vignette:
    code_folding: show
    toc: true
    toc_depth: 2
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Dave's plotting (and other) functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 4,
  fig.align = "center",
  dpi = 96
)
```

This vignette aims to show the main functionality of the functions that I (Dave) have developed around survey data summaries - doing those summaries, using the data and dictionary from those summaries to prepare survey data for further analysis/exploration, making plots in the same style (and, coming later, making other kinds of plots using the same theme and/or colour palettes).

Readers who just want to see how the plotting functions work should have a quick look at Section 1.2 to get some sense of the example/dummy data used in this vignette and then skip to Section 4 or later. 

# Setup

## Packages for this vignette

Load the `OMESurvey` package and a couple of other packages (inoffensive, commonly-used ones hopefully) that will be useful.
```{r setup}
library(OMESurvey)
library(tibble)
library(dplyr)
library(ggplot2)
```


## Some dummy data

Now create some dummy survey data and a corresponding data dictionary to use in the vignette. The data contains the following variables:

- `respondent_id`: a unique respondent identifier (included in the raw data but deliberately omitted from the dictionary to demonstrate handling of extra variables).
- `StartDate`: a survey start date (also omitted from the dictionary so that extra_vars behaviour can be illustrated).
- `school_type`: a categorical grouping variable with values Type A and Type B.
- `gender`: a second categorical grouping variable.
- `q_satisfaction`, `q_belonging`, and `q_support`: three Likert-style attitude questions with the same response scale (to demonstrate categorical summaries and `summary_plot_stacked_bar()`).
- `study_hours`: a numeric survey response giving estimated weekly study hours.
- `wellbeing_score` and `confidence_score`: numeric score variables on a 0–10 scale (used to demonstrate `OME_boxplot()` and `summary_plot_boxplot()`).
- `q_welcome_support_helpful`: A Likert-style question about whether they found welcome week support helpful.
- `q_finance_support_used`, `q_finance_support_helpful`: a yes/no branching variable indicating whether the respondent used finance support services and, for those who answered Yes, a Likert style question about whether they found it helpful.
- `q_academic_support_used`, `q_academic_support_helpful`: as above for academic support.
- `q_other_comments`: a free-text response variable (used to demonstrate how text fields are handled in the report).


The dummy data has been set up to

- use the -888 branching sentinel/missing code as should be the case for real OME survey data,
- include some -999 missing values,
- include a few -777 invalid values, an out-of-range numeric value, and a non-allowed categorical response.

These are included so that the validation and diagnostic parts of the report have something meaningful to show and the examples of plotting functions can show their capabilities.

```{r make_dummy_data_and_dict, echo=FALSE}
set.seed(123)

n <- 80

likert_levels <- c(
  "Strongly disagree",
  "Disagree",
  "Neither",
  "Agree",
  "Strongly agree"
)


helpful_levels <- c(
  "Very helpful",
  "Somewhat helpful",
  "Not very helpful"
)



# 1. Dummy survey data

well_conf_rho <- 0.3
well_conf_temp <-
  MASS::mvrnorm(
    n,
    mu = c(5, 4),
    Sigma = matrix(
      c(2^2, well_conf_rho*2*2.5,
        well_conf_rho*2*2.5, 2.5^2),
      nrow = 2
    )
  )

example_data <- tibble(
  respondent_id = sprintf("R%03d", seq_len(n)),  # deliberately not in dictionary
  StartDate = as.Date("2026-01-01") + sample(0:30, n, replace = TRUE), # extra var

  school_type = sample(c("Type A", "Type B"), n, replace = TRUE),
  gender = sample(
    c("Female", "Male", "Another gender", "Prefer not to say"),
    n,
    replace = TRUE,
    prob = c(0.45, 0.45, 0.04, 0.06)
  ),

  q_satisfaction = sample(likert_levels, n, replace = TRUE,
                          prob = c(0.08, 0.12, 0.20, 0.35, 0.25)),
  q_belonging = sample(likert_levels, n, replace = TRUE,
                       prob = c(0.10, 0.15, 0.25, 0.30, 0.20)),
  q_support = sample(likert_levels, n, replace = TRUE,
                     prob = c(0.06, 0.14, 0.20, 0.35, 0.25)),

  study_hours = round(pmax(0, rnorm(n, mean = 12, sd = 5)), 1),
  wellbeing_score = pmin(10, pmax(0, round(well_conf_temp[, 1]))),
  confidence_score = pmin(10, pmax(0, round(well_conf_temp[, 2]))),


q_finance_support_used = sample(
  c("Yes", "No"),
  n,
  replace = TRUE,
  prob = c(0.30, 0.70)
),
q_finance_support_helpful = NA_character_,

q_academic_support_used = sample(
  c("Yes", "No"),
  n,
  replace = TRUE,
  prob = c(0.45, 0.55)
),
q_academic_support_helpful = NA_character_,

q_welcome_support_helpful = sample(
  c(helpful_levels, "-999"),
  n,
  replace = TRUE,
  prob = c(0.42, 0.36, 0.14, 0.08)
),

q_other_comments = sample(
  c(
    "More worked examples would help",
    "I like the sessions",
    "Timetabling is difficult",
    "More advance notice would be useful",
    "-999"
  ),
  n,
  replace = TRUE,
  prob = c(0.2, 0.2, 0.2, 0.2, 0.2)
)

)

# Fill routed finance support follow-up:
# If q_finance_support_used == "Yes", provide a helpfulness response.
# Otherwise use -888 = not branched.
example_data <- example_data |>
  dplyr::mutate(
    q_finance_support_helpful = dplyr::if_else(
      q_finance_support_used == "Yes",
      sample(
        c(helpful_levels, "-999"),
        dplyr::n(),
        replace = TRUE,
        prob = c(0.40, 0.35, 0.15, 0.10)
      ),
      "-888"
    )
  )

# Fill routed academic support follow-up:
# If q_academic_support_used == "Yes", provide a helpfulness response.
# Otherwise use -888 = not branched.
example_data <- example_data |>
  dplyr::mutate(
    q_academic_support_helpful = dplyr::if_else(
      q_academic_support_used == "Yes",
      sample(
        c(helpful_levels, "-999"),
        dplyr::n(),
        replace = TRUE,
        prob = c(0.45, 0.35, 0.12, 0.08)
      ),
      "-888"
    )
  )



# Add a few deliberate edge cases

# General categorical examples
example_data$q_support[3] <- "Not sure"       # non-allowed factor value
example_data$q_belonging[7] <- "-999"         # explicit missing
example_data$q_satisfaction[10] <- "-777"     # invalid response

# Numeric examples
example_data$study_hours[4] <- 45             # out of allowed numeric range
example_data$wellbeing_score[8] <- -999       # explicit numeric missing
example_data$confidence_score[12] <- -777     # invalid numeric sentinel

# Branching variable sentinel examples
example_data$q_finance_support_used[15] <- "-999"  # missing branching variable
example_data$q_finance_support_helpful[15] <- "-999"       # follow-up also missing/unknown

example_data$q_academic_support_used[20] <- "-777"      # invalid branching variable
example_data$q_academic_support_helpful[20] <- "-999"      # follow-up missing/unknown

# Branched/follow-up question sentinel and non-allowed examples
example_data$q_finance_support_helpful[25] <- "-999"       # missing routed follow-up
example_data$q_academic_support_helpful[30] <- "-777"      # invalid routed follow-up

# Unrouted helpfulness question examples
example_data$q_welcome_support_helpful[45] <- "-999"       # missing response
example_data$q_welcome_support_helpful[50] <- "-777"       # invalid response



# Make the data dictionary

example_dictionary <- tibble(
  variable_name = c(
    "school_type",
    "gender",
    "q_satisfaction",
    "q_belonging",
    "q_support",
    "study_hours",
    "wellbeing_score",
    "confidence_score",
    
    "q_finance_support_used",
    "q_finance_support_helpful",
    "q_academic_support_used",
    "q_academic_support_helpful",
    "q_welcome_support_helpful",
    "q_other_comments"
  ),

  data_type = c(
    "factor-unordered",
    "factor-unordered",
    "factor-neg-pos",
    "factor-neg-pos",
    "factor-neg-pos",
    "numeric",
    "numeric",
    "numeric",

    "factor-unordered",
    "factor-lo-hi",
    "factor-unordered",
    "factor-lo-hi",
    "factor-lo-hi",
    "text"
  ),

  allowed_values = c(
    "Type A; Type B",
    "Female; Male; Another gender; Prefer not to say",
    paste(likert_levels, collapse = "; "),
    paste(likert_levels, collapse = "; "),
    paste(likert_levels, collapse = "; "),
    "[0,40]",
    "[0,10]",
    "[0,10]",

    "Yes; No",
    paste(helpful_levels, collapse = "; "),
    "Yes; No",
    paste(helpful_levels, collapse = "; "),
    paste(helpful_levels, collapse = "; "),
    NA_character_

  ),

  item_statement = c(
    "School type",
    "Gender",
    "I am satisfied with the programme",
    "I feel that I belong on the programme",
    "I know where to get support if I need it",
    "How many hours do you usually study outside sessions each week?",
    "Overall wellbeing score",
    "Confidence score",

    "Have you used financial support this term?",
    "How helpful was the financial support?",
    "Have you used academic support this term?",
    "How helpful was the academic support?",
    "How helpful was welcome week support?",

    "Please add any other comments about the programme"
    ),

  report_sec = c(
    "Grouping variables",
    "Grouping variables",
    "A: Attitudes",
    "A: Attitudes",
    "A: Attitudes",
    "B: Study habits",
    "B: Study habits",
    "B: Study habits",

    "C: Support use",
    "C: Support use",
    "C: Support use",
    "C: Support use",
    "C: Support use",
    "D: Free-text comments"
  ),


  grouping_var = c(
    TRUE,
    TRUE,
    FALSE,
    FALSE,
    FALSE,
    FALSE,
    FALSE,
    FALSE,

    FALSE,
    FALSE,
    FALSE,
    FALSE,
    FALSE,
    FALSE
  ),

  condition = c(
    NA_character_,
    NA_character_,
    NA_character_,
    NA_character_,
    NA_character_,
    NA_character_,
    NA_character_,
    NA_character_,

    NA_character_,
    'q_finance_support_used == "Yes"',
    NA_character_,
    'q_academic_support_used == "Yes"',
    NA_character_,
    NA_character_
  )
)
```



## Preview the example inputs

Before rendering a report or preparing the data, it is useful to briefly inspect both the raw survey data and the dictionary.

Preparing the data and dictionary is part of pre-processing, so this should need doing once only and is something that readers of this document will not have to worry much about. Guidance on doing that is in the Preprocessing SOP (which, at time of writing, is the Word document `20260408_SOP_Pre-processing` in the/a folder of SOPs). This includes information about constructing the data dictionary and what the -999/-888/-777 missing codes mean / are used for. An exception to this might be if you are doing a detailed data analysis (say, for a paper) and want to take a copy of the data dictionary to edit so that you can automatically prepare the data in a way that's useful for your analysis (rather than in a way that's useful for summarising the whole survey, which is the intention of the main data dictionary)

```{r preview-data, echo=FALSE}
if (requireNamespace("reactable", quietly = TRUE)) {

  reactable::reactable(
    example_data,
    wrap = TRUE,
    showPageSizeOptions = TRUE,
    pageSizeOptions = c(5, 10, 25, 50, 100),
    defaultPageSize = 5,
    fullWidth = FALSE,
    striped = TRUE,
    compact = TRUE,
    filterable = TRUE,
    resizable = TRUE
  )
} else {
  dplyr::glimpse(example_data)
}
```

```{r preview-dictionary, echo=FALSE}

if (requireNamespace("reactable", quietly = TRUE)) {

  example_dictionary |>
    dplyr::mutate(
      dplyr::across(
        dplyr::where(is.character),
        ~ stringr::str_replace_all(.x, "\\r?\\n", "  ")
      )
    ) |>
    reactable::reactable(
      wrap = TRUE,
      showPageSizeOptions = TRUE,
      pageSizeOptions = c(5, 10, 25, 50, 100),
      defaultPageSize = 5,
      fullWidth = FALSE,
      striped = TRUE,
      compact = TRUE,
      filterable = TRUE,
      resizable = TRUE
    )
} else {
  dplyr::glimpse(example_dictionary)
}
```


## Save example inputs to temporary files

For the purposes of the vignette, we save the dummy data as a CSV file and the corresponding dictionary as an Excel workbook. The dictionary is written to a sheet called "example_survey", which is the sheet name we will pass to `render_survey_summary()` and `survey_prepare_data()` below.

```{r save_data_and_dict, echo=FALSE}
data_path <- tempfile(pattern = "example_data_", fileext = ".csv")
dict_path <- tempfile(pattern = "example_survey_dictionary_", fileext = ".xlsx")

readr::write_csv(example_data, data_path)

openxlsx::write.xlsx(
  x = list(example_survey = example_dictionary),
  file = dict_path,
  overwrite = TRUE
)

data_path
dict_path
```


# Automated summary report

One can now use `render_survey_summary()` to create an HTML summary report from the dummy survey data and dictionary, using code along the lines of the following. This isn't done here as it's tricky to implement in a vignette and it's not the main point of this document. (There's more info in the pre-processing SOP.)

```{r render_report, eval=FALSE}
render_survey_summary(
  data_path = <path to data file>,
  dict_path = <path to dictionary file>,
  dict_sheet = "example_survey",
  output_file = <filename for output file, e.g. 'year_X_summary.html'>,
  output_dir = <path to folder for output file>,
  output_title = "Year X pupil survey summary",
  output_author = "A Nonymous"
)
```

<!-- The returned value of this function is the path to the generated HTML file. Open it in a browser to view the automated report. -->


# Prepare the data for further analysis

The report function first prepares the data internally (turning a csv/xlsx data file into a better-structured R data frame, e.g. ensuring that factors have the right levels) and then uses that data frame (and other validation information based on the data and data dictionary) to build the report. But it seems likely to be useful to have the data preparation phase separated, for further analysis and plotting.

The function `survey_prepare_data()` can be used to read the same csv/Excel data and Excel data dictionary, then validate/coerce the dictionary-backed variables to the appropriate numeric/factor types. The function returns a list containing several objects: most important for our purposes are the data (`data`) and a data frame that contains lots of validation information (`validation_df`). (There are also lists of messages that arose during the data & dictionary reading and processing, a list version of the validation information and other objects concerning the treatment of variables in the source data but not in the dictionary.)

```{r width=60}
prep <- survey_prepare_data(
  data_path = data_path,
  dict_path = dict_path,
  dict_sheet = "example_survey"
)

survey_data <- prep$data
validation_df <- prep$validation_df
```

In order to further customise the data this produces, one can make and edit a copy of the data dictionary. This allows full control over which variables are/aren't included, which values are treated as allowed/valid, etc. (One can also have establishment characteristics data merged into the survey data - see the SOP for info on how to do that, for now at least.)

We can take a quick look at the prepared data; noting particularly the names of the variables: those in the dictionary have the raw character-based version of the variable with the `_raw` suffix and the original variable name containing the tidied-and-coerced-to-type variable, and those not mentioned in the dictionary are appended with `_asis` (emphasising that they've not be processed in any way, in accordance with the default for the `extra_vars` option for `survey_prepare_data()`).


```{r echo=FALSE}
if (requireNamespace("reactable", quietly = TRUE)) {
  reactable::reactable(
    survey_data,
    wrap = TRUE,
    showPageSizeOptions = TRUE,
    pageSizeOptions = c(5, 10, 25, 50, 100),
    defaultPageSize = 5,
    fullWidth = FALSE,
    striped = TRUE,
    compact = TRUE,
    filterable = TRUE,
    resizable = TRUE
  )
} else {
  dplyr::glimpse(survey_data)

}
```

The validation output summarises the checks carried out for each dictionary-backed variable. (Here we show only some of the key columns in the `validation_df` output.)

```{r echo=FALSE}
if (requireNamespace("reactable", quietly = TRUE)) {
  validation_df |>
    dplyr::select(
      variable,
      type,
      allowed_specified,
      allowed_display,
      condition,
      report_sec
    ) |>
    reactable::reactable(
      wrap = FALSE,
      showPageSizeOptions = TRUE,
      pageSizeOptions = c(10, 25, 50, 100),
      defaultPageSize = 25,
      fullWidth = FALSE,
      striped = TRUE,
      compact = TRUE,
      filterable = TRUE,
      resizable = TRUE,
      columns = list(
        variable = reactable::colDef(minWidth = 200)
        # type = colDef(minWidth = 100),
        # allowed_specified = colDef(minWidth = 100),
        # allowed_display = colDef(minWidth = 100),
        # condition = colDef(minWidth = 100),
        # report_sec = colDef(minWidth = 100)
      )
    )
} else {
  validation_df |>
    dplyr::select(
      variable,
      type,
      allowed_specified,
      allowed_display,
      condition,
      report_sec
    )
}
```

# Categorical variables: stacked bar charts

## Single-question categorical plot with `OME_stacked_bar()`

`OME_stacked_bar()` is used for plotting one categorical survey response, optionally broken down by a grouping variable. Here we show satisfaction by school type

```{r}
OME_stacked_bar(
  dat = survey_data,
  response_var = q_satisfaction,
  group_var = school_type,
  count_style = "both",
  titleText = "Satisfaction by year group"
)
```

The default fill colours for the bars are (the 'distinct' or 'qualitative' palette type from `OMESurvey::get_OME_colours()`), designed to be as distinct/easy-to-distinguish as possible. This palette can be modified manually by passing, for example, `OMESurvey::get_OME_colours(n=5, type='divergent')` as the `colo` argument of the plotting function. But the validation output already has that set up (from the info in the data dictionary) and ready to use:

```{r}
validation_df |> filter(variable=="q_satisfaction") |> pull(colo) |> (\(x) x[[1]])()
```

(Aside: the `(\(x) x[[1]])()` at the end of the code extracts the first item from a list - it is needed because of how the validation info is stored, which in turn is needed to avoid some technical issues that result from what R tries to do when you include a vector as an element of a vector. The colo variable in validation_df is a list-column, with entries either NA or a vector describing the palette.)

Passing this to `OME_stacked_bar()` we get
```{r}
OME_stacked_bar(
  dat = survey_data,
  response_var = q_satisfaction,
  group_var = school_type,
  colo = validation_df |> filter(variable=="q_satisfaction") |> pull(colo) |> (\(x) x[[1]])(),
  count_style = "both",
  titleText = "Satisfaction by year group"
)
```

This is a common recurring theme for factors and stacked bar charts: you need to tell the plotting function what colour palette to use, but the appropriate one is always available through the validation data frame:
```{r eval=FALSE}
validation_df |> filter(variable=="appropriate_variable_name") |> pull(colo) |> (\(x) x[[1]])()
```

(This method is preferable to passing `OMESurvey::get_OME_colours(n=5, type='divergent')` to the plotting function's `colo` argument because it returns a named vector, which makes the matching of levels and colours less likely to go wrong in some unlikely-but-realistic edge-case scenarios.)




We can use the same method to look at the distribution of responses to another Likert-scale type question, here (i) grouped by a different variable and (ii) with missing/invalid values excluded from the stacked bars (but still indicated in the numbers in parentheses).

```{r}
OME_stacked_bar(
  dat = survey_data,
  response_var = q_support,
  group_var = gender,
  na.rm = TRUE,
  colo = validation_df |> filter(variable=="q_support") |> pull(colo) |> (\(x) x[[1]])(),
  count_style = "both",
  titleText = "Knowing where to get support, by gender"
)
```


### An example with routing taken account of

It is fairly simple to deal with routing here directly: filter the data appropriately before passing it to the plotting function.

```{r}
survey_data |>
  filter(q_finance_support_used == "Yes") |>
  OME_stacked_bar(
    response_var = q_finance_support_helpful,
    colo = validation_df |>
      filter(variable == "q_finance_support_helpful") |>
      pull(colo) |>
      (\(x) x[[1]])(),
    na.rm = TRUE,
    count_style = "both",
    titleText = "Helpfulness of financial support"
  )
```

(It's possible to do this in a slightly more programmatic way, using information about routing from `validation_df`, but for now I'm leaving that out.)


## Multi-question categorical plot with `summary_plot_stacked_bar()`

`summary_plot_stacked_bar()` summarises several categorical questions in one horizontal stacked bar chart. The variables representing the questions must share the same response scale, like the three attitude questions in the dummy data, for example.

(In principle one could `pivot_longer()` the data for the relevant variables manually and then use `OME_stacked_bar()` as above. But (i) this is a common task and (ii) it gets tricky when dealing with missing values and/or survey branching conditions. Hence the specially-written function.)

```{r}
attitude_vars <- c("q_satisfaction", "q_belonging", "q_support")

# extract labels that match vars
attitude_labels <- validation_df |>
  dplyr::filter(variable %in% attitude_vars) |>
  dplyr::arrange(match(variable, attitude_vars)) |>
  dplyr::pull(item_statement) |>
  stats::setNames(attitude_vars)

# extract colour palette
attitude_colo <- validation_df |>
  dplyr::filter(variable == attitude_vars[1]) |>
  dplyr::pull(colo) |>
  (\(x) x[[1]])()

survey_data |>
  dplyr::select(q_satisfaction, q_belonging, q_support) |>
  summary_plot_stacked_bar(
    colo = attitude_colo,
    labels_vec = attitude_labels,
    count_style = "both",
    titleText = "Attitudes to the programme"
  )
```

A couple of things to note here:

- The vector/palette of colours `attitude_colo` is simply taken from one of the questions. (I could have something fancier, but it would have been hard to code. You need to make sure that the variables you are using have the same allowed values / response options!)
- We need to `select()` the variables to use in the data frame and then pass that to the plotting function. The function treats every variable in the data frame it receives as a question to be plotted, hence the need to select() them first. Or, of course, use a base R equivalent: `survey_data[,c("q_satisfaction", "q_belonging", "q_support")] |> summary_plot_stacked_bar(...)`.


### Ordering questions

The questions can also be ordered according to a response pattern using the `order_values` option. For example, we can order by the proportion of respondents answering `"Agree"` or `"Strongly agree"`. In this example we also show how to extract some more details (the question labels as well as the colour palette) from the `validation_df` rather than hard-coding them.

```{r}
attitude_vars <- c("q_satisfaction", "q_belonging", "q_support")

# extract labels that match vars
attitude_labels <- validation_df |>
  dplyr::filter(variable %in% attitude_vars) |>
  dplyr::arrange(match(variable, attitude_vars)) |>
  dplyr::pull(item_statement) |>
  stats::setNames(attitude_vars)

# extract colour palette
attitude_colo <- validation_df |>
  dplyr::filter(variable == attitude_vars[1]) |>
  dplyr::pull(colo) |>
  (\(x) x[[1]])() 

# plot
survey_data |>
  dplyr::select(dplyr::all_of(attitude_vars)) |>
  summary_plot_stacked_bar(
    colo = attitude_colo,
    labels_vec = attitude_labels,
    count_style = "both",
    order_values = c("Agree", "Strongly agree"),
    titleText = "Attitudes ordered by positive responses"
  )
```


### Dealing with survey routing

**It's a bit awkward**

The three helpfulness questions use the same response scale, but we do not have data for the same respondents across all the questions. Financial-support helpfulness is only relevant to respondents who used financial support, academic-support helpfulness is only relevant to respondents who used academic support, while the welcome-week helpfulness question is asked of everyone. Just plotting these variables is misleading as we do not respect the routing of the survey (i.e. the conditions in the data dictionary).


```{r}
helpful_vars <- c(
  "q_finance_support_helpful",
  "q_academic_support_helpful",
  "q_welcome_support_helpful"
)

# Extract labels and colours from validation_df
helpful_labels <-
  validation_df |>
  dplyr::filter(variable %in% helpful_vars) |>
  dplyr::arrange(match(variable, helpful_vars)) |>
  dplyr::pull(item_statement) |>
  stats::setNames(helpful_vars)

helpful_colo <-
  validation_df |>
  dplyr::filter(variable == helpful_vars[1]) |>
  dplyr::pull(colo) |>
  (\(x) x[[1]])()

# plot
survey_data |>
  dplyr::select(dplyr::all_of(helpful_vars)) |>
  summary_plot_stacked_bar(
    labels_vec = helpful_labels,
    colo = helpful_colo,
    count_style = "both",
    titleText = "Helpfulness of support",
  )
```

As the following table makes explicit, only 25 respondents were asked about the helpfulness of financial support. And then, only 20 of them gave a substantive response. The label "(20/80)" and the 75% missing bar/label do not convey this at all.


```{r}
survey_data |>
  janitor::tabyl(q_finance_support_helpful, q_finance_support_used) |>
  janitor::adorn_totals(where="both") |>
  janitor::adorn_title() |>
  kable_narrow()
```



Simply removing missing values (using the `na.rm=TRUE` option) is an improvement, and together with some explanation it might be good enough for some purposes (especially if we tidy up the labels by using the `count_style="non-missing"` argument or even omit them using `show_counts=FALSE`). But it still doesn't convey all the information that we have. While most missing / non-allowed values are because of routing, several are not; neither the labels nor the plot properly capture that there are 20 substantive responses, 5 missing/invalid responses, 55 non-routed non-responses.


**But there's a way around it**

The `summary_plot_stacked_bar()` function supports an “extended” data format to deal with this, where each plotted variable has a name ending `_value` and companion variables ending `_include`, and `_plot` which together capture the substantive/missing/irrelevant structure:

- `*_value`: the response value to plot;
- `*_include`: whether the respondent should be included in the denominator of that question's label;
- `*_plot`: whether the response is a valid substantive value to show (i.e. should be included in the plot and numerator of the label).

Of course it's possible to construct these `*_value`, `*_include`, `*_plot` variables yourself, using what you know about the routing conditions for each variable. But it's fiddly. So there's a function `make_extended_summary_plot_data()` to do it automatically, using the information from the data dictionary, which is contained in the data/dictionary validation information (which we've already used to get colour palettes & survey item labels). Provide the function with the data and validation information from `survey_prepare_data()` and the variables you want to work with and it gives you a dataframe ready for passing to `summary_plot_stacked_bar()`.

Here's how we can use it for the helpfulness questions in our dummy survey.

```{r}
helpful_ext <- make_extended_summary_plot_data(
  data = survey_data,
  validation_df = validation_df,
  vars = helpful_vars
)
```

Then we can use the same `summary_plot_stacked_bar()` function with this extended-format data.
```{r}
helpful_ext |>
  summary_plot_stacked_bar(
    dat_format = "extended",
    labels_vec = helpful_labels,
    colo = helpful_colo,
    na.rm = TRUE,
    count_style = "both",
    titleText = "Helpfulness of support"
  )

```

Note especially the label "(20/25)" for the financial support question, capturing the 20 substantive responses (and 5 missing/invalid/non-allowed responses) amongst the 25 respondents who were routed to this question. The same is true for the academic support question, as could be verified by constructing a crosstab of the `q_academic_support_used` and `q_academic_support_helpful` variables. And the welcome week support question, which was asked to everyone, still has the full number of respondents indicated (reflecting it being asked to everyone and thus not having a condition in the data dictionary).

**Including the missing/invalid responses**

If you do want to include the missing/invalid responses in the stacked bars, you can do so by changing the `na.rm` option to `FALSE`. This will ensure that the missing responses are included in the bars, but still shows the number of valid responses and number of routed responses in the annotation on the right (this can be varied using the `count_style` option).

```{r}
helpful_ext |>
  summary_plot_stacked_bar(
    dat_format = "extended",
    labels_vec = helpful_labels,
    colo = helpful_colo,
    na.rm = FALSE,
    count_style = "both",
    titleText = "Helpfulness of support"
  )
```

## Other tweaking for stacked bar charts

Unless otherwise stated, these methods are applicable to plots created with both `OME_stacked_bar()` and `summary_plot_stacked_bar()`.

### Font size

There is a `base_size` option that works for both `OME_stacked_bar()` and `summary_plot_stacked_bar()`. The base size is given in points and is the size of the title text, with axis/legend title text 0.85 times the base size and axis/legend label text 0.7 times that size. The default 14 makes these sizes just under 12 and 11 points, respectively. But for a presentation one might want to make all the text a bit bigger:
```{r}
helpful_ext |>
  summary_plot_stacked_bar(
    dat_format = "extended",
    labels_vec = helpful_labels,
    colo = helpful_colo,
    base_size = 18
  )
```


### Label text wrapping

If labels are particularly long then changing the way they are wrapped can be helpful to make the plot easier to read. This is more likely to be an issue with larger text size too.

The `group_label_width` option is the target number of characters used before breaking to a new line (with `NULL` meaning 'no line breaks'). The defaults are `r deparse(formals(OME_boxplot_)$group_label_width)` for `OME_boxplot()` and `r deparse(formals(summary_plot_boxplot)$group_label_width)` for `summary_plot_boxplot()`. For example:
```{r}
helpful_ext |>
  summary_plot_stacked_bar(
    dat_format = "extended",
    labels_vec = helpful_labels,
    colo = helpful_colo,
    group_label_width = 20
  )
```

And `fill_label_width` does the same for the legend labels (with default 20).
```{r}
helpful_ext |>
  summary_plot_stacked_bar(
    dat_format = "extended",
    labels_vec = helpful_labels,
    colo = helpful_colo,
    fill_label_width = 10
  )
```

### Legend positioning/alignment

There is another helpful trick available if the legend is below the plot (i.e. spread out horizontally) and needs a bit more space: use the `legend.location` option of the ggplot2 theme. The default is 'panel', which aligns the legend with the plotting panel only, but the alternative 'plot' aligns it with the whole area of the figure:
```{r}
helpful_ext |>
  summary_plot_stacked_bar(
    dat_format = "extended",
    labels_vec = helpful_labels,
    colo = helpful_colo
    ) +
  theme(legend.location = 'plot')
```



# Numerical variables: boxplots

## Single-question numeric plot with `OME_boxplot()`

`OME_boxplot()` is the numeric equivalent of `OME_stacked_bar()`. It plots one numeric variable, optionally broken down by a factor grouping variable. The main features and options are very similar to those in `OME_boxplot()`.

```{r}
OME_boxplot(
  data = survey_data,
  value_var = study_hours,
  group_var = school_type,
  titleText = "Weekly study hours by school type"
)
```

As with stacked bars, the data can be filtered before being passed to the plotting function (to allow for routing or simply to focus on a particular subset of respondents). Here I show how to use the colour argument to the function to control the colour of the boxplot.

```{r}
survey_data |>
  filter(gender=="Female") |>
OME_boxplot(
  value_var = study_hours,
  group_var = school_type,
  titleText = "Weekly study hours by school type",
  colour = "#2244AA"
)
```


## Multi-question numeric plot with `summary_plot_boxplot()`

`summary_plot_boxplot()` summarises several numeric questions in one horizontal boxplot. Here we compare study hours, wellbeing, and confidence.

```{r}
numeric_labels <- c(
  wellbeing_score = "Wellbeing score",
  confidence_score = "Confidence score"
)

survey_data |>
  dplyr::select(study_hours, wellbeing_score, confidence_score) |>
  summary_plot_boxplot(
    labels_vec = numeric_labels,
    titleText = "Numeric survey measures"
  )
```

### Ordering of questions

By default, the questions are ordered by their median value. To keep the original variable order instead, set `order_fun = NULL`. (`mean` also works here.)

```{r}
survey_data |>
  dplyr::select(wellbeing_score, confidence_score) |>
  summary_plot_boxplot(
    labels_vec = numeric_labels,
    order_fun = NULL,
    titleText = "Numeric survey measures in original order"
  )
```


### Dealing with survey routing

This is essentially the same as for the stacked bar charts. Use `make_extended_summary_plot_data()` to take the survey data and set up `*_value`, `*_include` and `*_plot` columns for each variable, then `summary_plot_boxplot()` to plot.

## Other tweaking for boxplots

Unless otherwise stated, these methods are applicable to plots created with both `OME_boxplot()` and `summary_plot_boxplot()`.

### Font size

The `base_size` option works for boxplots too:
```{r}
survey_data |>
  OME_boxplot(
    study_hours,
    school_type,
    base_size = 18
  )
```

### Label text wrapping

If labels are particularly long then changing the way they are wrapped can be helpful to make the plot easier to read. This is more likely to be an issue with larger text size too.

The `group_label_width` option controls how many characters wide the grouping labels should be. For example:
```{r}
survey_data |>
  dplyr::select(wellbeing_score, confidence_score) |>
  summary_plot_boxplot(
    labels_vec = numeric_labels,
    order_fun = NULL,
    group_label_width = 10
  )
```

### Setting scale limits and/or guide values

If the numerical variables being plotted are percentages then it might make sense to insist that the scale runs from 0 to 100 and that the guide lines should be at multiples of 25. Here's an example where manually specifying the scale doesn't make much sense, but nevertheless shows how it can be done. (Note that using just one of `coord_cartesian` and `scale_x_continuous` will work too.)
```{r}
survey_data |>
  OME_boxplot(
    study_hours,
    school_type
  ) +
  coord_cartesian(xlim = c(0, 50)) +
  scale_x_continuous(breaks = seq(0, 50, by = 12.5))
```





# A generic theme and colour palettes

For other kinds of plots, I have made

- a ggplot2 theme that should make plots in an OME kind of way, and
- fill & colour scales that use the get_OME_colours() colour palettes.

## Generic theme `theme_OME()`

If I make some other kind of plot and want it to be OME-ey, then I can apply the theme `OME_theme()`.

```{r}
survey_data |>
  ggplot(aes(x = confidence_score,
             y = wellbeing_score)) +
  geom_point(position = 'jitter',
             alpha = 0.6) +
  labs(x = "Confidence score",
       y = "Wellbeing score") +
  coord_cartesian(xlim=c(0,10), ylim=c(0,10)) +
  theme_OME()
```


And if I use colour then I can set the colour scale to be on-brand too. (This scale also encourages the OME-ey style by removing the title in the legend, by default.)
```{r}
survey_data |>
  ggplot(aes(x = confidence_score,
             y = wellbeing_score,
             colour = school_type)) +
  geom_point(position = 'jitter',
             alpha = 0.6) +
  labs(x = "Confidence score",
       y = "Wellbeing score") +
  coord_cartesian(xlim=c(0,10), ylim=c(0,10)) +
  theme_OME() +
  scale_colour_OME()
```

It defaults to using the 'distinct' palette, but the other options ('divergent'/'sequential') can work too, e.g. using `scale_colour_OME(type='sequential')`.
```{r}
survey_data |>
  ggplot(aes(x = confidence_score,
             y = wellbeing_score,
             colour = school_type)) +
  geom_point(position = 'jitter',
             alpha = 0.8) +
  labs(x = "Confidence score",
       y = "Wellbeing score") +
  coord_cartesian(xlim=c(0,10), ylim=c(0,10)) +
  theme_OME() +
  scale_colour_OME(type='sequential')
```


And the same sort of thing can be achieved for the `fill` aesthetic:
```{r}
survey_data |>
  ggplot(aes(x = school_type,
             fill = gender)) +
  geom_bar() +
  labs(x = NULL,
       y = "Count") +
  theme_OME() +
  theme(panel.grid.major.x = element_blank()) +
  scale_fill_OME()
```


# Other useful things

... more to come here soon, including

- using `r `forcats::fct_reorder()` to reorder factors
- using patchwork to put plots together (and consolidating legends)
- saving advice
- other things too maybe...


# Still coming...


There will also, in due course, be additional functions to do other common(ish)ly needed types of plots.


```{r}
knitr::knit_exit()
```



# Using `theme_OME()`, `scale_colour_OME()`, and `scale_fill_OME()`

The OME theme and colour scales can also be used directly with ordinary `ggplot2` plots.

```{r}
survey_data |>
  ggplot(aes(
    x = study_hours,
    y = wellbeing_score,
    colour = school_type
  ) ) +
  geom_point(size = 2, alpha = 0.8) +
  scale_colour_OME() +
  theme_OME() +
  labs(
    title = "Wellbeing and study hours",
    x = "Weekly study hours",
    y = "Wellbeing score",
    colour = "School type"
  )
```

The fill scale works similarly for geoms that have a `fill`. Here we use a proportional bar chart to show satisfaction by school type

```{r}
survey_data |>
  ggplot2::ggplot(
    ggplot2::aes(
      x = school_type,
      fill = q_satisfaction
    )
  ) +
  ggplot2::geom_bar(position = "fill") +
  scale_fill_OME() +
  theme_OME(base_family = "sans") +
  ggplot2::labs(
    title = "Satisfaction by school type",
    x = "Year group",
    y = "Proportion of responses",
    fill = "Satisfaction"
  )
```

(Brief note/aside: the US-spelling alias `scale_color_OME()` can be used in place of `scale_colour_OME()`.)



## Extra variables and naming conventions

Because we prepared the data with `extra_vars = "suffix_asis"`, variables present in the source data but absent from the dictionary are retained with an `_asis` suffix.

```{r}
names(survey_data)[grepl("_asis$", names(survey_data))]
```

Dictionary-backed variables also have `_raw` companions. These preserve the raw values used during validation and condition checking.

```{r}
names(survey_data)[grepl("_raw$", names(survey_data))]
```

For example, we can compare a prepared factor variable with its raw version.

```{r}
survey_data |>
  dplyr::select(q_support, q_support_raw) |>
  dplyr::distinct() |>
  dplyr::arrange(q_support_raw)
```

The raw variable reveals the deliberately inserted non-allowed value `"Not sure"`, while the prepared variable is coerced according to the dictionary-defined allowed values.

## Text responses

The dummy data includes a free-text field. In the automated report, text variables are shown as an interactive table rather than plotted. For direct analysis, the prepared data can be filtered to remove sentinel values.

```{r}
survey_data |>
  dplyr::filter(!(q_other_comments %in% c("-999", "-888", "-777"))) |>
  dplyr::select(school_type, gender, q_other_comments) |>
  dplyr::slice_head(n = 10)
```

If `reactable` is installed, we can show the text responses interactively.

```{r echo=FALSE}
if (requireNamespace("reactable", quietly = TRUE)) {
  survey_data |>
    dplyr::filter(!(q_other_comments %in% c("-999", "-888", "-777"))) |>
    dplyr::select(school_type, gender, q_other_comments) |>
    reactable::reactable(
      wrap = TRUE,
      showPageSizeOptions = TRUE,
      pageSizeOptions = c(10, 25, 50, 100),
      defaultPageSize = 10,
      fullWidth = FALSE,
      striped = TRUE,
      compact = TRUE,
      filterable = TRUE,
      resizable = TRUE
    )
}
```

## Cleaning up temporary files

The example files were written to temporary locations. They will usually be removed automatically at the end of the R session, but they can also be deleted manually.

```{r}
#unlink(c(data_path, dict_path, report_path))
```
