---
title: "Preparation and initial survey analysis"
---

```{r, include = FALSE, eval = F}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
requireNamespace("tibble", quietly = TRUE)
requireNamespace("DT", quietly = TRUE)

```

```{r setup}
library(OMESurvey)
```

*This article shows how to use tools from * **OMESurvey** *to check and clean the provided survey data and produce an initial report on 'missingness' in the survey*


In this article we will show how to prepare and perform initial survey analysis on fake data based on the year 7 pilot study. This fake data has the same structure as those we receive. We show the data for 5 participants (transposed so the table is easier to see).

```{r}
data = OMESurvey::survey_example

# Remove the themes and add spacing to imagine how the data will arrive)
split = names(data) |> stringr::str_split('__')
names(data) = lapply(split, function(x){if(length(x) == 1){return(x)} ; x[2]}) |> unlist() |> stringr::str_replace_all('_', ' ')
```
```{r, echo = F}
data[1:5,] |>  t() |> DT::datatable(extensions = 'Buttons',
                  rownames = T,
                  options = list(scrollY = 300,
                                 dom = 'Bt',
                                 buttons = c('copy', 'csv', 'excel', 'pdf'),
                                 paging=FALSE,
                                 ordering=F))
```

### Data Cleaning

Our first aim is to clean the data such that the data is ready to be analysed and no further modifications are required. Note that after the first clean we will need to do sanity checks in case there are other issues in the data. 

The initial part of cleaning will involve: 

- Adding theme / group information to the column names. To using the tools herein the column name of questions should be in the format `XX__YY` where XX is the theme/group and YY the question (i.e we use '__' double underscore to represent the split between group/theme and the question).
- Convert 'N-A' to NA. 
- Get the school ID from the bar code.
- Create a combined school and class ID (in case multiple schools have the same class ID).

```{r}
# Clean 1. 

# A) Adding theme/group to column names.
group = c('Theme 1' |> rep(6),
          'Theme 2'|> rep(7), 
          'Theme 3'|> rep(12), 
          'Theme 4'|> rep(4),
          'Theme 5'|> rep(12), 
          'Theme 6'|> rep(7),
          'Theme 7'|> rep(6), 
          'Theme 8'|> rep(7))             

names(data)[6:ncol(data)] = paste0(group, '__', names(data)[6:ncol(data)]) 
names(data) = names(data) |> stringr::str_replace_all(pattern = ' ', replacement = '_')

# B) NAs
for(i in 1:ncol(data)){
  data[,i][data[,i] == 'N-A'] = NA
  # data[,i] = data[,i] |> stringr::str_to_sentence() # commented out for example.
}

# C + D) School ID and combine school and class information.
school = data$Barcode_ID |> as.character() |>  substr(3, 5)
school_and_class = paste0(school,'__', data$`Group_/_class`) 

# Combine into a cleaned dataset. 
data_clean = data.frame(data[,1:4], school, school_and_class, data[,5:ncol(data)])
```

We show the data for same 5 participants as before (transposed so the table is easier to see).
```{r, echo = F}
data_clean[1:5,] |> t() |> DT::datatable(extensions = 'Buttons',
                  rownames = T,
                  options = list(scrollY = 300,
                                 dom = 'Bt',
                                 buttons = c('copy', 'csv', 'excel', 'pdf'),
                                 paging=FALSE,
                                 ordering=F))

```

After the first step of cleaning we see the addition of the school and school_class combined columns new column names and NAs added.


After the first step of cleaning we can now run checks for other errors in the dataset. 
To check for issues we can use `check_survey()` which checks for barcode issues, unexpected values in question columns and questions which have multiple answers (i.e. a value of "99").

The function returns a list informing if we have any issues of the type listed above. 

```{r}
check = OMESurvey::check_survey(data)
check
```


We see that most errors occur from casing issues (capital N in never), since this isn't case specific to this survey we can fix this using `stringr::str_to_sentence()` in the cleaning section (which is already there just commented out to show the issue). The issue of 'Neither' being used instead of 'Half the time' is case specific and can be fixed for this individual report. 

```{r}
# Run extra cleaning with issues found.
for(i in 1:ncol(data)){
  data_clean[,i] = data_clean[,i] |> stringr::str_to_sentence() 
}

# Fix neither to Half the time.
data_clean$Theme_3__Question_21[data_clean$Theme_3__Question_21 == 'Neither'] = 'Half the time'

```


If we want to check the survey images for records where we have "99s" we can filter the SharePoint search by the barcodes using the following text

```{r}
paste0(check$multiple_answers$barcodes, collapse = " OR ")
```


#### Checking for issues in the ID

As a further precaution we may want to check that the IDs we have in our survey dataset aligns with the information already held for ID and name combinations. This will be useful to track issues in the ID, changes in names or to check for children who used an 'empty form' rather then the survey meant for them (is it was missing or misplaced for example).

Within **OMESurvey** the function `check_IDs()` performs the ID checks. An example is shown below.

```{r}
# Use a large extract of the data as the 'true' values
true_df = data[1:100,]

# Small subset for our survey results.
data_ex  = data[1:10,]
#Change some names and IDs.
data_ex$First_Name[3] = 'Luke'
data_ex$Barcode_ID[7] = data_ex$Barcode_ID[7] |> stringr::str_extract(pattern = '[0-9]{8}')


# Check the IDs using 
check_IDs(data = data_ex, true_df = true_df)
```

***

### Missingness report

Now we have cleaned the data we can create a missingness report to explore if there are empty surveys, which questions are more often missing, etc. 

Within **OMESurvey** we have a pre-written markdown report which takes survey data as an input and analyses the missingness. Therefore we only need to run the function `create_missingness_report()` to create the report for whichever survey data we are analysing. 

```{r, eval = F}
# Create the missingness report.
OMESurvey::create_missingness_report(survey_data = data_clean, 
                                     survey_name = 'example survey',
                                     output_dir = getwd()
                                     )
```

Note that by default this creates a .html report where the charts are interactive (i.e has hover enabled, plus potential other features). This allows for more informative charts however it comes at the cost of not been able to edit the document. Therefore using the input `report_kind = 'static'` will render the same report except the output will be in .docx allowing text to be edited/added at the expense of the interactive plots (where they will be replaced with static screengrabs).
