---
title: "Create survey figures"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(OMESurvey)
```

*This vignette provides examples for the simple creation of survey plots across different output types*

```{r load some data}
data = OMESurvey::survey_example

# Clean such that "N-A" values equal NA, fix any casing issues
for(i in 1:ncol(data)){
  data[,i][data[,i] == 'N-A'] = NA
  data[,i] = data[,i] |> stringr::str_to_sentence() 
}

```

## Number of responses across questions in a theme

In this case we can use `plot_theme()` to create a horizontal bar chart. Depending on the "Kind" argument we can create a ggplot, a plotly figure or a formatted data frame.

### {.tabset}

#### ggplot

```{r plot theme - ggplot}
OMESurvey::plot_theme(data, theme = 'Theme_1', kind = 'ggplot')
```

#### plotly

```{r plot theme - plotly}
OMESurvey::plot_theme(data, theme = 'Theme_1', kind = 'plotly')
```

#### data.frame

```{r plot theme data frame}
df = OMESurvey::plot_theme(data, theme = 'Theme_1', kind = 'data.frame')
df |> DT::datatable(extensions = 'Buttons',
                  rownames = F,
                  options = list(scrollY = 300,
                                 dom = 'Bt',
                                 buttons = c('copy', 'csv', 'excel', 'pdf'),
                                 paging=FALSE,
                                 ordering=F))
```

### {-}


Note that here we use default `survey_values = OMESurvey::survey_values` is used to find the expected values of the survey questions and the order to show the answers in the legend. For custom survey values you can change the `survey_values`. The function uses colours from `OMESurvey::get_OME_colours()`. For the plots either a ggplot or plotly object are returned, therefore + or |> can be used to update the plot (e.g layout/colours, etc). Or you can create your own plot from the data frame output.

For example if you want to swap the bar order but keep the same colourscheme.

```{r, plot theme - ggplot 2}
# Use custom survey_values with the order of the values reversed
p = OMESurvey::plot_theme(data, theme = 'Theme_1', kind = 'ggplot',
                      survey_values = list(agree = c("Agree a lot", "Agree a little", "Neither",           "Disagree a little", "Disagree a lot") |> rev()))

# Get OME colours
colo = get_OME_colours(n = 5, type = 'contrast') 
# Swap to the original colour order
p + ggplot2::scale_fill_manual(values = colo |> rev(), guide = ggplot2::guide_legend(reverse = TRUE)) 
```

Note further a warning will be given if the answers to the questions do not align with the survey values. Where the unexpected answers will be set to NA (for ggplot).

```{r, plot theme - ggplot 3, warning=T}
data_issue = data
data_issue$Theme_1__Question_1[1] = 'Ohh Dear...'

# Use custom survey_values with the order of the values reversed
OMESurvey::plot_theme(data_issue, theme = 'Theme_1', kind = 'ggplot')

```

***

## Number of responses across questions in a theme with a demographic

Need to add a demographic column so extract school from the bar code.

```{r}
school = data$Barcode_ID |> as.character() |>  substr(3, 5)
data$school = school
school_column = ncol(data)
```
### {.tabset}

#### ggplot

```{r plot theme demo - ggplot}
OMESurvey::plot_theme_by_demographic(data, theme = 'Theme_1', demographic_column = school_column,  kind = 'ggplot')
```


#### plotly

```{r plot theme demo - plotly}
OMESurvey::plot_theme_by_demographic(data, theme = 'Theme_1', demographic_column = school_column,  kind = 'plotly')
```

#### data.frame

```{r plot theme demo data frame}
# Get school information
df = OMESurvey::plot_theme_by_demographic(data, theme = 'Theme_1', demographic_column = school_column,  kind = 'data.frame')

df |> DT::datatable(extensions = 'Buttons',
                  rownames = F,
                  options = list(scrollY = 300,
                                 dom = 'Bt',
                                 buttons = c('copy', 'csv', 'excel', 'pdf'),
                                 paging=FALSE,
                                 ordering=F))
```

### {-}

## Adding OME logo to ggplot

You can also add the OME logo to ggplot objects using `add_logo()`.

Note however to do this we use the gridExtra::grid.arrange() function to add the logo into the margin which causes the object to no longer be a ggplot object. Thus, you cannot use '+' to change the plot after adding the logo.

```{r}
# Add logo.
OMESurvey::plot_theme(data, theme = 'Theme_1', kind = 'ggplot') |> OMESurvey::add_logo()

# Add logo, speciify position.
OMESurvey::plot_theme(data, theme = 'Theme_1', kind = 'ggplot') |> OMESurvey::add_logo(position = 'top right')

# Add logo, speciify position and sizing arguement.
OMESurvey::plot_theme(data, theme = 'Theme_1', kind = 'ggplot') |> OMESurvey::add_logo(position = 'top left',logo_sizing = c(0.3, 0.5))
```

## Formatted bar charts from data column

We can use `plot_bar()` to create bar charts of a single column within survey data. 

### {.tabset}

#### ggplot

```{r plot bar - ggplot}
OMESurvey::plot_bar(data, column = 5, kind = 'ggplot')
```

#### plotly

```{r plot bar - plotly}
OMESurvey::plot_bar(data, column = 5, kind = 'plotly')
```

#### data.frame

```{r plot bar data frame}
df = OMESurvey::plot_bar(data, column = 5, kind = 'data.frame')
df |> DT::datatable(extensions = 'Buttons',
                  rownames = F,
                  options = list(scrollY = 300,
                                 dom = 'Bt',
                                 buttons = c('copy', 'csv', 'excel', 'pdf'),
                                 paging=FALSE,
                                 ordering=F))
```

### {-}

Note that you can include NA in the bar charts by using passing arguments to `table()` via `...`. Moreover, for ggplot outputs we can include numbers on the plot y setting `showValue = TRUE`.

```{r}
OMESurvey::plot_bar(data, column = 5, kind = 'ggplot', showValue = TRUE, useNA = 'always')
```

